### 2020-03-07 acknowledge.tex
tweaking stuff

\mbox{}\\
\noindent
\textbf{Acknowledgments:} The authors thank John Glasser for comments
and discussion on the manuscript, particularly on short notice. 
This work was supported, in part,
by support from the Army Research Office (W911NF1910384).

\noindent
\textbf{Data Availabililty:} All code is available at https://github.com/mac-theobio/coronavirus_asymptomatic.