### 2020-03-07 appendix.tex

\section*{Supplementary Materials}
\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\theequation{S\arabic{equation}}
\setcounter{figure}{0}
\setcounter{equation}{0}

\subsection*{A compartmental model for asymptomatic/symptomatic cases}

Consider an SEIR model variant in which an infected individual can be either asymptomatic, $I_a$, or symptomatic, $I_s$.
While both cases can recover, we assume that only symptomatic cases can lead to fatalities, denoted
by the $D$ category.
In total, the dynamics of susceptibles, exposed, infectious, recovered, and dead are:
\begin{eqnarray}
\dot{S}&=&-\beta_a S I_a -\beta_s S I_s \\
\dot{E}&=&\beta_a S I_a +\beta_s S I_s -\gamma_e E\\
\dot{I}_a&=&p\gamma_e E-\gamma_a I_a\\
\dot{I}_s&=&(1-p)\gamma_e E-\gamma_s I_s\\
\dot{R}&=&\gamma_a I_a + (1-f)\gamma_s I_s \\
\dot{D}&=&f\gamma_s I_s.
\end{eqnarray}
Here, $\beta_a$ and $\beta_s$ denote transmission rates,
$\gamma_e$ denotes the transition from exposed to infectious,
$p$ is the fraction of asymptomatic cases that
are generated for each exposed individual,
$1-p$ is the fraction of symptomatic cases that
are generated for each exposed individual,
$\gamma_a$ and $\gamma_s$ denote recovery rates,
and $f$ denotes the case fatality ratio for symptomatic cases.

Given that the number of infected individuals increase exponentially at rate $r$ initially, the equations
for the infectious cases can be rewritten given the ansatz
$E(t)=c_e e^{rt}$,
$I_a(t)=c_a e^{rt}$,
$I_s(t)=c_s e^{rt}$.
Then, it follows that 
\begin{eqnarray}
r c_a&=&p\gamma_e c_e-\gamma_a c_a,\\
r c_s&=&(1-p)\gamma_e c_e-\gamma_s c_s,
\end{eqnarray} 
which implies that
\begin{equation}
\frac{c_a}{c_s}=\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right].
\end{equation}
Therefore, the ratio of secondary case production caused by asymptomatic vs.~symptomatic
individuals during the exponential phase should be
\begin{equation}
\frac{q}{1-q} = \left(\frac{\beta_a}{\beta_s}\right)\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right],
\end{equation}
where $q$ is the fraction of new secondary cases caused by
asymptomatic individuals.

The basic reproduction number of this system is:
\begin{equation}
{\cal{R}}_{0}= p {\cal{R}}_a + (1-p) {\cal{R}}_s,
\end{equation}
where 
\begin{eqnarray}
{\cal{R}}_a &=& \frac{\beta_a}{\gamma_a},\\
{\cal{R}}_s &=& \frac{\beta_s}{\gamma_s}.
\end{eqnarray}
The generation-interval distributions for asymptomatic and symptomatic individuals follow the same functional form as the corresponding generation-interval distribution for a single-type SEIR model since both asymptomatic and symptomatic individuals have exponenitally distributed latent and infectious periods \citep{svensson2015influence}:
\begin{eqnarray}
g_a(\tau) &=& \frac{\gamma_e \gamma_a}{\gamma_e - \gamma_a} \left(\exp(-\gamma_a \tau) - \exp(-\gamma_e \tau)\right),\\
g_s(\tau) &=& \frac{\gamma_e \gamma_s}{\gamma_e - \gamma_s} \left(\exp(-\gamma_s \tau) - \exp(-\gamma_e \tau)\right).
\end{eqnarray}
It immediately follows that 
\begin{equation}
\left(\frac{z}{1-z}\right)\left[\frac{\int_0^\infty \exp(-r\tau) g_a(\tau) \mathrm{d}\tau}{\int_0^\infty \exp(-r\tau) g_s(\tau) \mathrm{d}\tau}\right] = \left(\frac{\beta_a}{\beta_s}\right)\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right],
\end{equation}
where $z = p {\cal{R}}_a/{\cal{R}}_{0}$ and $1-z=(1-p) {\cal{R}}_s/{\cal{R}}_{0}$.

\pagebreak

\subsection*{Supplementary figures}

\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap_02.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0_02.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases when generation-interval distributions are narrow. 
See Figure 1 in the main text for figure caption.
Here, we assume $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=\kappa_a=0.2$.
}
\end{center}
\end{figure*}

\pagebreak

\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap_08.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0_08.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases when generation-interval distributions are wide. 
See Figure 1 in the main text for figure caption.
Here, we assume $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=\kappa_a=0.8$.
}
\end{center}
\end{figure*}

\pagebreak

\begin{figure*}[!ht]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_kappa_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the generation-interval dispersion of asymptomatic cases.
(A) Wide/narrow generation intervals of asymptomatic cases increase/decrease the relevance of asymptomatic cases, $q$.
(B) Wide/narrow generation intervals of asymptomatic cases decrease/increase the basic reproduction number ${\cal{R}}_0$.
As our baseline parameters, we assume $1/r=7$ days, $\bar G_s=G_a=7$ days, and $\kappa_s=0.5$.}
\end{center}
\end{figure*}