### 2020-03-09 figheatmap_R0.tex
push all files

% Created by tikzDevice version 0.12.3 on 2020-03-09 11:51:36
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (397.48,325.21);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (397.48,325.21);
\end{scope}
\begin{scope}
\path[clip] ( 34.16, 30.69) rectangle (338.79,302.56);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.16, 30.69) rectangle (338.79,302.56);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 44.28) --
	(338.79, 44.28);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 71.47) --
	(338.79, 71.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 98.65) --
	(338.79, 98.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,125.84) --
	(338.79,125.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,153.03) --
	(338.79,153.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,180.22) --
	(338.79,180.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,207.40) --
	(338.79,207.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,234.59) --
	(338.79,234.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,261.78) --
	(338.79,261.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,288.96) --
	(338.79,288.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] (110.31, 30.69) --
	(110.31,302.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] (262.63, 30.69) --
	(262.63,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	(338.79, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 57.87) --
	(338.79, 57.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 85.06) --
	(338.79, 85.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,112.25) --
	(338.79,112.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,139.43) --
	(338.79,139.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,166.62) --
	(338.79,166.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,193.81) --
	(338.79,193.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,221.00) --
	(338.79,221.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,248.18) --
	(338.79,248.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,275.37) --
	(338.79,275.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,302.56) --
	(338.79,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	( 34.16,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (186.47, 30.69) --
	(186.47,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.79, 30.69) --
	(338.79,302.56);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at ( 35.25,  32.05) {
	\pgfimage[width=303.25pt,height=269.15pt,interpolate=false]{figheatmap_R0_ras1}};
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,192.10) --
	( 35.87,193.81) --
	( 36.35,194.30) --
	( 38.51,196.53) --
	( 38.55,196.57) --
	( 40.75,198.92) --
	( 41.05,199.25) --
	( 42.95,201.34) --
	( 43.50,201.97) --
	( 45.14,203.83) --
	( 45.87,204.68) --
	( 47.34,206.41) --
	( 48.16,207.40) --
	( 49.54,209.08) --
	( 50.38,210.12) --
	( 51.74,211.84) --
	( 52.52,212.84) --
	( 53.93,214.69) --
	( 54.59,215.56) --
	( 56.13,217.64) --
	( 56.60,218.28) --
	( 58.33,220.69) --
	( 58.54,221.00) --
	( 60.43,223.71) --
	( 60.53,223.85) --
	( 62.26,226.43) --
	( 62.72,227.14) --
	( 64.03,229.15) --
	( 64.92,230.54) --
	( 65.76,231.87) --
	( 67.12,234.07) --
	( 67.43,234.59) --
	( 69.06,237.31) --
	( 69.32,237.74) --
	( 70.64,240.03) --
	( 71.51,241.56) --
	( 72.18,242.75) --
	( 73.68,245.46) --
	( 73.71,245.53) --
	( 75.13,248.18) --
	( 75.91,249.66) --
	( 76.55,250.90) --
	( 77.94,253.62) --
	( 78.11,253.96) --
	( 79.28,256.34) --
	( 80.30,258.44) --
	( 80.60,259.06) --
	( 81.88,261.78) --
	( 82.50,263.12) --
	( 83.13,264.50) --
	( 84.35,267.21) --
	( 84.70,268.01) --
	( 85.54,269.93) --
	( 86.70,272.65) --
	( 86.90,273.11) --
	( 87.83,275.37) --
	( 88.95,278.09) --
	( 89.09,278.45) --
	( 90.03,280.81) --
	( 91.09,283.53) --
	( 91.29,284.04) --
	( 92.12,286.25) --
	( 93.14,288.96) --
	( 93.49,289.91) --
	( 94.13,291.68) --
	( 95.10,294.40) --
	( 95.69,296.06) --
	( 96.05,297.12) --
	( 96.98,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 90.39) --
	( 34.45, 90.50) --
	( 36.35, 91.21) --
	( 38.55, 92.05) --
	( 40.75, 92.91) --
	( 41.50, 93.22) --
	( 42.95, 93.81) --
	( 45.14, 94.73) --
	( 47.34, 95.69) --
	( 47.90, 95.94) --
	( 49.54, 96.67) --
	( 51.74, 97.69) --
	( 53.75, 98.65) --
	( 53.93, 98.75) --
	( 56.13, 99.84) --
	( 58.33,100.97) --
	( 59.10,101.37) --
	( 60.53,102.14) --
	( 62.72,103.35) --
	( 64.02,104.09) --
	( 64.92,104.61) --
	( 67.12,105.92) --
	( 68.57,106.81) --
	( 69.32,107.28) --
	( 71.51,108.69) --
	( 72.78,109.53) --
	( 73.71,110.15) --
	( 75.91,111.68) --
	( 76.70,112.25) --
	( 78.11,113.27) --
	( 80.30,114.93) --
	( 80.35,114.97) --
	( 82.50,116.66) --
	( 83.76,117.68) --
	( 84.70,118.47) --
	( 86.90,120.36) --
	( 86.95,120.40) --
	( 89.09,122.33) --
	( 89.94,123.12) --
	( 91.29,124.40) --
	( 92.76,125.84) --
	( 93.49,126.57) --
	( 95.42,128.56) --
	( 95.69,128.84) --
	( 97.88,131.23) --
	( 97.92,131.28) --
	(100.08,133.75) --
	(100.29,134.00) --
	(102.28,136.40) --
	(102.53,136.72) --
	(104.47,139.19) --
	(104.66,139.43) --
	(106.67,142.14) --
	(106.69,142.15) --
	(108.61,144.87) --
	(108.87,145.25) --
	(110.44,147.59) --
	(111.07,148.56) --
	(112.18,150.31) --
	(113.26,152.06) --
	(113.85,153.03) --
	(115.44,155.75) --
	(115.46,155.78) --
	(116.97,158.47) --
	(117.66,159.74) --
	(118.43,161.18) --
	(119.83,163.90) --
	(119.86,163.97) --
	(121.17,166.62) --
	(122.05,168.48) --
	(122.45,169.34) --
	(123.69,172.06) --
	(124.25,173.32) --
	(124.88,174.78) --
	(126.03,177.50) --
	(126.45,178.52) --
	(127.13,180.22) --
	(128.20,182.93) --
	(128.65,184.12) --
	(129.22,185.65) --
	(130.21,188.37) --
	(130.84,190.16) --
	(131.17,191.09) --
	(132.09,193.81) --
	(132.99,196.53) --
	(133.04,196.70) --
	(133.85,199.25) --
	(134.69,201.97) --
	(135.24,203.80) --
	(135.50,204.68) --
	(136.28,207.40) --
	(137.04,210.12) --
	(137.44,211.55) --
	(137.78,212.84) --
	(138.50,215.56) --
	(139.20,218.28) --
	(139.63,220.03) --
	(139.87,221.00) --
	(140.53,223.71) --
	(141.17,226.43) --
	(141.79,229.15) --
	(141.83,229.35) --
	(142.39,231.87) --
	(142.98,234.59) --
	(143.55,237.31) --
	(144.03,239.63) --
	(144.11,240.03) --
	(144.65,242.75) --
	(145.18,245.46) --
	(145.69,248.18) --
	(146.20,250.90) --
	(146.23,251.05) --
	(146.69,253.62) --
	(147.17,256.34) --
	(147.63,259.06) --
	(148.09,261.78) --
	(148.42,263.79) --
	(148.54,264.50) --
	(148.97,267.21) --
	(149.40,269.93) --
	(149.81,272.65) --
	(150.22,275.37) --
	(150.62,278.09) --
	(150.62,278.09) --
	(151.01,280.81) --
	(151.39,283.53) --
	(151.76,286.25) --
	(152.13,288.96) --
	(152.48,291.68) --
	(152.82,294.27) --
	(152.84,294.40) --
	(153.18,297.12) --
	(153.51,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41, 48.53) --
	(335.21, 48.76) --
	(333.01, 48.99) --
	(330.81, 49.24) --
	(328.62, 49.49) --
	(326.69, 49.72) --
	(326.42, 49.75) --
	(324.22, 50.02) --
	(322.02, 50.29) --
	(319.83, 50.58) --
	(317.63, 50.87) --
	(315.43, 51.18) --
	(313.23, 51.50) --
	(311.04, 51.83) --
	(308.84, 52.17) --
	(307.21, 52.44) --
	(306.64, 52.53) --
	(304.44, 52.90) --
	(302.25, 53.28) --
	(300.05, 53.68) --
	(297.85, 54.09) --
	(295.65, 54.53) --
	(293.46, 54.98) --
	(292.63, 55.15) --
	(291.26, 55.45) --
	(289.06, 55.94) --
	(286.86, 56.45) --
	(284.67, 56.99) --
	(282.47, 57.55) --
	(281.27, 57.87) --
	(280.27, 58.14) --
	(278.07, 58.76) --
	(275.88, 59.41) --
	(273.68, 60.09) --
	(272.15, 60.59) --
	(271.48, 60.81) --
	(269.28, 61.57) --
	(267.09, 62.37) --
	(264.89, 63.22) --
	(264.66, 63.31) --
	(262.69, 64.11) --
	(260.49, 65.06) --
	(258.39, 66.03) --
	(258.30, 66.07) --
	(256.10, 67.14) --
	(253.90, 68.29) --
	(253.06, 68.75) --
	(251.70, 69.51) --
	(249.51, 70.82) --
	(248.47, 71.47) --
	(247.31, 72.22) --
	(245.11, 73.73) --
	(244.48, 74.19) --
	(242.92, 75.36) --
	(240.98, 76.90) --
	(240.72, 77.12) --
	(238.52, 79.03) --
	(237.88, 79.62) --
	(236.32, 81.11) --
	(235.11, 82.34) --
	(234.13, 83.39) --
	(232.63, 85.06) --
	(231.93, 85.88) --
	(230.39, 87.78) --
	(229.73, 88.63) --
	(228.36, 90.50) --
	(227.53, 91.68) --
	(226.51, 93.22) --
	(225.34, 95.07) --
	(224.81, 95.94) --
	(223.26, 98.65) --
	(223.14, 98.87) --
	(221.82,101.37) --
	(220.94,103.16) --
	(220.50,104.09) --
	(219.27,106.81) --
	(218.74,108.03) --
	(218.12,109.53) --
	(217.06,112.25) --
	(216.55,113.62) --
	(216.06,114.97) --
	(215.13,117.68) --
	(214.35,120.09) --
	(214.25,120.40) --
	(213.42,123.12) --
	(212.65,125.84) --
	(212.15,127.67) --
	(211.91,128.56) --
	(211.22,131.28) --
	(210.56,134.00) --
	(209.95,136.67) --
	(209.94,136.72) --
	(209.35,139.43) --
	(208.79,142.15) --
	(208.25,144.87) --
	(207.76,147.53) --
	(207.75,147.59) --
	(207.26,150.31) --
	(206.79,153.03) --
	(206.35,155.75) --
	(205.92,158.47) --
	(205.56,160.90) --
	(205.52,161.18) --
	(205.12,163.90) --
	(204.75,166.62) --
	(204.39,169.34) --
	(204.04,172.06) --
	(203.71,174.78) --
	(203.39,177.50) --
	(203.36,177.75) --
	(203.08,180.22) --
	(202.78,182.93) --
	(202.49,185.65) --
	(202.22,188.37) --
	(201.95,191.09) --
	(201.69,193.81) --
	(201.44,196.53) --
	(201.20,199.25) --
	(201.16,199.64) --
	(200.96,201.97) --
	(200.73,204.68) --
	(200.51,207.40) --
	(200.30,210.12) --
	(200.09,212.84) --
	(199.89,215.56) --
	(199.70,218.28) --
	(199.51,221.00) --
	(199.32,223.71) --
	(199.14,226.43) --
	(198.97,229.15) --
	(198.97,229.24) --
	(198.80,231.87) --
	(198.64,234.59) --
	(198.47,237.31) --
	(198.32,240.03) --
	(198.17,242.75) --
	(198.02,245.46) --
	(197.87,248.18) --
	(197.73,250.90) --
	(197.59,253.62) --
	(197.46,256.34) --
	(197.33,259.06) --
	(197.20,261.78) --
	(197.08,264.50) --
	(196.95,267.21) --
	(196.83,269.93) --
	(196.77,271.48) --
	(196.72,272.65) --
	(196.60,275.37) --
	(196.49,278.09) --
	(196.38,280.81) --
	(196.27,283.53) --
	(196.17,286.25) --
	(196.07,288.96) --
	(195.97,291.68) --
	(195.87,294.40) --
	(195.77,297.12) --
	(195.68,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,112.42) --
	(335.21,113.46) --
	(333.01,114.54) --
	(332.16,114.97) --
	(330.81,115.65) --
	(328.62,116.79) --
	(326.96,117.68) --
	(326.42,117.98) --
	(324.22,119.20) --
	(322.14,120.40) --
	(322.02,120.47) --
	(319.83,121.78) --
	(317.66,123.12) --
	(317.63,123.14) --
	(315.43,124.54) --
	(313.48,125.84) --
	(313.23,126.00) --
	(311.04,127.51) --
	(309.56,128.56) --
	(308.84,129.08) --
	(306.64,130.71) --
	(305.89,131.28) --
	(304.44,132.40) --
	(302.44,134.00) --
	(302.25,134.16) --
	(300.05,135.98) --
	(299.19,136.72) --
	(297.85,137.88) --
	(296.13,139.43) --
	(295.65,139.87) --
	(293.46,141.93) --
	(293.23,142.15) --
	(291.26,144.09) --
	(290.48,144.87) --
	(289.06,146.34) --
	(287.88,147.59) --
	(286.86,148.69) --
	(285.41,150.31) --
	(284.67,151.15) --
	(283.06,153.03) --
	(282.47,153.73) --
	(280.82,155.75) --
	(280.27,156.43) --
	(278.68,158.47) --
	(278.07,159.26) --
	(276.64,161.18) --
	(275.88,162.23) --
	(274.69,163.90) --
	(273.68,165.36) --
	(272.83,166.62) --
	(271.48,168.66) --
	(271.04,169.34) --
	(269.33,172.06) --
	(269.28,172.13) --
	(267.68,174.78) --
	(267.09,175.79) --
	(266.10,177.50) --
	(264.89,179.67) --
	(264.59,180.22) --
	(263.13,182.93) --
	(262.69,183.77) --
	(261.72,185.65) --
	(260.49,188.11) --
	(260.37,188.37) --
	(259.06,191.09) --
	(258.30,192.73) --
	(257.80,193.81) --
	(256.59,196.53) --
	(256.10,197.64) --
	(255.41,199.25) --
	(254.28,201.97) --
	(253.90,202.88) --
	(253.18,204.68) --
	(252.12,207.40) --
	(251.70,208.48) --
	(251.09,210.12) --
	(250.09,212.84) --
	(249.51,214.47) --
	(249.12,215.56) --
	(248.18,218.28) --
	(247.31,220.90) --
	(247.28,221.00) --
	(246.39,223.71) --
	(245.54,226.43) --
	(245.11,227.81) --
	(244.70,229.15) --
	(243.89,231.87) --
	(243.11,234.59) --
	(242.92,235.26) --
	(242.34,237.31) --
	(241.59,240.03) --
	(240.87,242.75) --
	(240.72,243.33) --
	(240.16,245.46) --
	(239.47,248.18) --
	(238.81,250.90) --
	(238.52,252.08) --
	(238.15,253.62) --
	(237.51,256.34) --
	(236.89,259.06) --
	(236.32,261.61) --
	(236.29,261.78) --
	(235.69,264.50) --
	(235.11,267.21) --
	(234.55,269.93) --
	(234.13,272.03) --
	(234.00,272.65) --
	(233.46,275.37) --
	(232.93,278.09) --
	(232.42,280.81) --
	(231.93,283.46) --
	(231.92,283.53) --
	(231.42,286.25) --
	(230.94,288.96) --
	(230.46,291.68) --
	(230.00,294.40) --
	(229.73,296.05) --
	(229.55,297.12) --
	(229.11,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,163.53) --
	(336.92,163.90) --
	(335.21,165.22) --
	(333.45,166.62) --
	(333.01,166.97) --
	(330.81,168.77) --
	(330.14,169.34) --
	(328.62,170.64) --
	(326.99,172.06) --
	(326.42,172.57) --
	(324.22,174.56) --
	(323.98,174.78) --
	(322.02,176.61) --
	(321.11,177.50) --
	(319.83,178.75) --
	(318.36,180.22) --
	(317.63,180.95) --
	(315.72,182.93) --
	(315.43,183.24) --
	(313.23,185.61) --
	(313.19,185.65) --
	(311.04,188.06) --
	(310.77,188.37) --
	(308.84,190.61) --
	(308.43,191.09) --
	(306.64,193.25) --
	(306.19,193.81) --
	(304.44,196.00) --
	(304.03,196.53) --
	(302.25,198.86) --
	(301.95,199.25) --
	(300.05,201.83) --
	(299.95,201.97) --
	(298.02,204.68) --
	(297.85,204.92) --
	(296.15,207.40) --
	(295.65,208.14) --
	(294.35,210.12) --
	(293.46,211.50) --
	(292.60,212.84) --
	(291.26,215.00) --
	(290.92,215.56) --
	(289.29,218.28) --
	(289.06,218.66) --
	(287.71,221.00) --
	(286.86,222.48) --
	(286.17,223.71) --
	(284.69,226.43) --
	(284.67,226.48) --
	(283.25,229.15) --
	(282.47,230.67) --
	(281.85,231.87) --
	(280.50,234.59) --
	(280.27,235.05) --
	(279.18,237.31) --
	(278.07,239.66) --
	(277.90,240.03) --
	(276.66,242.75) --
	(275.88,244.49) --
	(275.45,245.46) --
	(274.27,248.18) --
	(273.68,249.58) --
	(273.12,250.90) --
	(272.01,253.62) --
	(271.48,254.93) --
	(270.92,256.34) --
	(269.86,259.06) --
	(269.28,260.57) --
	(268.83,261.78) --
	(267.82,264.50) --
	(267.09,266.53) --
	(266.84,267.21) --
	(265.88,269.93) --
	(264.95,272.65) --
	(264.89,272.82) --
	(264.03,275.37) --
	(263.14,278.09) --
	(262.69,279.49) --
	(262.27,280.81) --
	(261.42,283.53) --
	(260.59,286.25) --
	(260.49,286.55) --
	(259.77,288.96) --
	(258.98,291.68) --
	(258.30,294.06) --
	(258.20,294.40) --
	(257.44,297.12) --
	(256.69,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,205.34) --
	(335.37,207.40) --
	(335.21,207.57) --
	(333.01,209.87) --
	(332.78,210.12) --
	(330.81,212.24) --
	(330.27,212.84) --
	(328.62,214.69) --
	(327.86,215.56) --
	(326.42,217.22) --
	(325.53,218.28) --
	(324.22,219.84) --
	(323.28,221.00) --
	(322.02,222.55) --
	(321.10,223.71) --
	(319.83,225.35) --
	(319.00,226.43) --
	(317.63,228.25) --
	(316.96,229.15) --
	(315.43,231.26) --
	(314.99,231.87) --
	(313.23,234.37) --
	(313.08,234.59) --
	(311.23,237.31) --
	(311.04,237.60) --
	(309.44,240.03) --
	(308.84,240.95) --
	(307.69,242.75) --
	(306.64,244.43) --
	(306.00,245.46) --
	(304.44,248.04) --
	(304.36,248.18) --
	(302.76,250.90) --
	(302.25,251.79) --
	(301.21,253.62) --
	(300.05,255.70) --
	(299.70,256.34) --
	(298.23,259.06) --
	(297.85,259.76) --
	(296.79,261.78) --
	(295.65,264.00) --
	(295.40,264.50) --
	(294.04,267.21) --
	(293.46,268.41) --
	(292.72,269.93) --
	(291.43,272.65) --
	(291.26,273.02) --
	(290.17,275.37) --
	(289.06,277.83) --
	(288.95,278.09) --
	(287.75,280.81) --
	(286.86,282.85) --
	(286.58,283.53) --
	(285.43,286.25) --
	(284.67,288.11) --
	(284.32,288.96) --
	(283.23,291.68) --
	(282.47,293.62) --
	(282.16,294.40) --
	(281.12,297.12) --
	(280.27,299.39) --
	(280.10,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,240.19) --
	(335.30,242.75) --
	(335.21,242.86) --
	(333.13,245.46) --
	(333.01,245.62) --
	(331.03,248.18) --
	(330.81,248.47) --
	(328.99,250.90) --
	(328.62,251.41) --
	(327.01,253.62) --
	(326.42,254.44) --
	(325.08,256.34) --
	(324.22,257.58) --
	(323.21,259.06) --
	(322.02,260.83) --
	(321.40,261.78) --
	(319.83,264.19) --
	(319.63,264.50) --
	(317.91,267.21) --
	(317.63,267.67) --
	(316.24,269.93) --
	(315.43,271.27) --
	(314.61,272.65) --
	(313.23,275.01) --
	(313.03,275.37) --
	(311.48,278.09) --
	(311.04,278.88) --
	(309.97,280.81) --
	(308.84,282.90) --
	(308.50,283.53) --
	(307.07,286.25) --
	(306.64,287.07) --
	(305.67,288.96) --
	(304.44,291.40) --
	(304.31,291.68) --
	(302.97,294.40) --
	(302.25,295.91) --
	(301.67,297.12) --
	(300.40,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,269.68) --
	(337.22,269.93) --
	(335.26,272.65) --
	(335.21,272.73) --
	(333.36,275.37) --
	(333.01,275.87) --
	(331.50,278.09) --
	(330.81,279.11) --
	(329.69,280.81) --
	(328.62,282.47) --
	(327.94,283.53) --
	(326.42,285.94) --
	(326.23,286.25) --
	(324.56,288.96) --
	(324.22,289.52) --
	(322.93,291.68) --
	(322.02,293.22) --
	(321.34,294.40) --
	(319.83,297.06) --
	(319.79,297.12) --
	(318.28,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,294.95) --
	(335.98,297.12) --
	(335.21,298.32) --
	(334.24,299.84);

\path[draw=drawColor,line width= 0.9pt,dash pattern=on 4pt off 4pt ,line join=round] (186.47, 30.69) -- (186.47,302.56);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (262.63, 35.04) {underestimated};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (110.31, 35.04) {overestimated};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.16, 30.69) rectangle (338.79,302.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 54.84) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 82.03) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,109.22) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,136.40) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,163.59) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,190.78) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,217.97) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,245.15) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,272.34) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,299.53) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 30.69) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 57.87) --
	( 34.16, 57.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 85.06) --
	( 34.16, 85.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,112.25) --
	( 34.16,112.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,139.43) --
	( 34.16,139.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,166.62) --
	( 34.16,166.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,193.81) --
	( 34.16,193.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,221.00) --
	( 34.16,221.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,248.18) --
	( 34.16,248.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,275.37) --
	( 34.16,275.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,302.56) --
	( 34.16,302.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 27.94) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (186.47, 27.94) --
	(186.47, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.79, 27.94) --
	(338.79, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 34.16, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (186.47, 19.68) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (338.79, 19.68) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (186.47,  7.64) {Relative mean asymptomatic generation interval, $G_a/G_s$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,166.62) {Intrinsic proportion of asymptomatic transmission, $z$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (349.79, 25.48) rectangle (391.98,307.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at (355.29,  30.98) {
	\pgfimage[width= 14.45pt,height=256.07pt,interpolate=true]{figheatmap_R0_ras2}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24, 34.35) {1.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24, 88.41) {2.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,142.47) {3.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,196.53) {4.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,250.59) {5.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (355.29,293.62) {$\mathcal{R}_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29, 37.38) -- (358.18, 37.38);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29, 91.44) -- (358.18, 91.44);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,145.50) -- (358.18,145.50);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,199.56) -- (358.18,199.56);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,253.62) -- (358.18,253.62);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85, 37.38) -- (369.74, 37.38);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85, 91.44) -- (369.74, 91.44);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,145.50) -- (369.74,145.50);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,199.56) -- (369.74,199.56);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,253.62) -- (369.74,253.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.16,310.62) {B. Basic reproduction number};
\end{scope}
\end{tikzpicture}

\end{document}