### 2020-03-09 figheatmap_R0_03.tex

% Created by tikzDevice version 0.12.3 on 2020-03-09 11:51:45
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (397.48,325.21);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (397.48,325.21);
\end{scope}
\begin{scope}
\path[clip] ( 34.16, 30.69) rectangle (338.79,302.56);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.16, 30.69) rectangle (338.79,302.56);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 44.28) --
	(338.79, 44.28);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 71.47) --
	(338.79, 71.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 98.65) --
	(338.79, 98.65);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,125.84) --
	(338.79,125.84);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,153.03) --
	(338.79,153.03);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,180.22) --
	(338.79,180.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,207.40) --
	(338.79,207.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,234.59) --
	(338.79,234.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,261.78) --
	(338.79,261.78);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,288.96) --
	(338.79,288.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] (110.31, 30.69) --
	(110.31,302.56);

\path[draw=drawColor,line width= 0.3pt,line join=round] (262.63, 30.69) --
	(262.63,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	(338.79, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 57.87) --
	(338.79, 57.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 85.06) --
	(338.79, 85.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,112.25) --
	(338.79,112.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,139.43) --
	(338.79,139.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,166.62) --
	(338.79,166.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,193.81) --
	(338.79,193.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,221.00) --
	(338.79,221.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,248.18) --
	(338.79,248.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,275.37) --
	(338.79,275.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,302.56) --
	(338.79,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	( 34.16,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (186.47, 30.69) --
	(186.47,302.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.79, 30.69) --
	(338.79,302.56);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at ( 35.25,  32.72) {
	\pgfimage[width=303.25pt,height=267.79pt,interpolate=false]{figheatmap_R0_03_ras1}};
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,189.02) --
	( 34.89,189.73) --
	( 36.28,191.09) --
	( 36.35,191.16) --
	( 37.64,192.45) --
	( 38.55,193.38) --
	( 38.97,193.81) --
	( 40.28,195.17) --
	( 40.75,195.66) --
	( 41.57,196.53) --
	( 42.83,197.89) --
	( 42.95,198.01) --
	( 44.07,199.25) --
	( 45.14,200.44) --
	( 45.29,200.61) --
	( 46.48,201.97) --
	( 47.34,202.95) --
	( 47.66,203.32) --
	( 48.81,204.68) --
	( 49.54,205.55) --
	( 49.95,206.04) --
	( 51.06,207.40) --
	( 51.74,208.23) --
	( 52.16,208.76) --
	( 53.24,210.12) --
	( 53.93,211.01) --
	( 54.30,211.48) --
	( 55.34,212.84) --
	( 56.13,213.88) --
	( 56.37,214.20) --
	( 57.37,215.56) --
	( 58.33,216.86) --
	( 58.37,216.92) --
	( 59.34,218.28) --
	( 60.31,219.64) --
	( 60.53,219.95) --
	( 61.25,221.00) --
	( 62.19,222.36) --
	( 62.72,223.14) --
	( 63.11,223.71) --
	( 64.01,225.07) --
	( 64.90,226.43) --
	( 64.92,226.46) --
	( 65.78,227.79) --
	( 66.64,229.15) --
	( 67.12,229.91) --
	( 67.49,230.51) --
	( 68.33,231.87) --
	( 69.16,233.23) --
	( 69.32,233.48) --
	( 69.98,234.59) --
	( 70.78,235.95) --
	( 71.51,237.20) --
	( 71.57,237.31) --
	( 72.35,238.67) --
	( 73.12,240.03) --
	( 73.71,241.07) --
	( 73.88,241.39) --
	( 74.63,242.75) --
	( 75.37,244.11) --
	( 75.91,245.10) --
	( 76.10,245.46) --
	( 76.82,246.82) --
	( 77.53,248.18) --
	( 78.11,249.29) --
	( 78.23,249.54) --
	( 78.92,250.90) --
	( 79.60,252.26) --
	( 80.28,253.62) --
	( 80.30,253.67) --
	( 80.94,254.98) --
	( 81.60,256.34) --
	( 82.25,257.70) --
	( 82.50,258.23) --
	( 82.89,259.06) --
	( 83.52,260.42) --
	( 84.14,261.78) --
	( 84.70,263.00) --
	( 84.76,263.14) --
	( 85.37,264.50) --
	( 85.97,265.86) --
	( 86.56,267.21) --
	( 86.90,267.98) --
	( 87.15,268.57) --
	( 87.73,269.93) --
	( 88.30,271.29) --
	( 88.87,272.65) --
	( 89.09,273.19) --
	( 89.43,274.01) --
	( 89.98,275.37) --
	( 90.53,276.73) --
	( 91.07,278.09) --
	( 91.29,278.65) --
	( 91.60,279.45) --
	( 92.13,280.81) --
	( 92.65,282.17) --
	( 93.17,283.53) --
	( 93.49,284.37) --
	( 93.68,284.89) --
	( 94.18,286.25) --
	( 94.68,287.60) --
	( 95.18,288.96) --
	( 95.67,290.32) --
	( 95.69,290.38) --
	( 96.15,291.68) --
	( 96.62,293.04) --
	( 97.10,294.40) --
	( 97.57,295.76) --
	( 97.88,296.69) --
	( 98.03,297.12) --
	( 98.48,298.48) --
	( 98.94,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 63.05) --
	( 35.49, 63.31) --
	( 36.35, 63.48) --
	( 38.55, 63.94) --
	( 40.75, 64.40) --
	( 41.98, 64.67) --
	( 42.95, 64.88) --
	( 45.14, 65.38) --
	( 47.34, 65.89) --
	( 47.91, 66.03) --
	( 49.54, 66.42) --
	( 51.74, 66.97) --
	( 53.35, 67.39) --
	( 53.93, 67.54) --
	( 56.13, 68.13) --
	( 58.33, 68.74) --
	( 58.37, 68.75) --
	( 60.53, 69.37) --
	( 62.72, 70.02) --
	( 63.01, 70.11) --
	( 64.92, 70.70) --
	( 67.12, 71.40) --
	( 67.32, 71.47) --
	( 69.32, 72.13) --
	( 71.32, 72.83) --
	( 71.51, 72.89) --
	( 73.71, 73.68) --
	( 75.06, 74.19) --
	( 75.91, 74.51) --
	( 78.11, 75.36) --
	( 78.55, 75.54) --
	( 80.30, 76.26) --
	( 81.83, 76.90) --
	( 82.50, 77.19) --
	( 84.70, 78.17) --
	( 84.91, 78.26) --
	( 86.90, 79.18) --
	( 87.81, 79.62) --
	( 89.09, 80.25) --
	( 90.55, 80.98) --
	( 91.29, 81.36) --
	( 93.13, 82.34) --
	( 93.49, 82.53) --
	( 95.58, 83.70) --
	( 95.69, 83.76) --
	( 97.88, 85.05) --
	( 97.90, 85.06) --
	(100.08, 86.41) --
	(100.10, 86.42) --
	(102.19, 87.78) --
	(102.28, 87.83) --
	(104.19, 89.14) --
	(104.47, 89.34) --
	(106.08, 90.50) --
	(106.67, 90.93) --
	(107.90, 91.86) --
	(108.87, 92.61) --
	(109.63, 93.22) --
	(111.07, 94.40) --
	(111.28, 94.58) --
	(112.87, 95.94) --
	(113.26, 96.29) --
	(114.38, 97.29) --
	(115.46, 98.30) --
	(115.84, 98.65) --
	(117.24,100.01) --
	(117.66,100.43) --
	(118.58,101.37) --
	(119.86,102.72) --
	(119.87,102.73) --
	(121.11,104.09) --
	(122.05,105.16) --
	(122.31,105.45) --
	(123.46,106.81) --
	(124.25,107.77) --
	(124.58,108.17) --
	(125.65,109.53) --
	(126.45,110.58) --
	(126.68,110.89) --
	(127.68,112.25) --
	(128.65,113.60) --
	(128.65,113.61) --
	(129.59,114.97) --
	(130.50,116.33) --
	(130.84,116.86) --
	(131.37,117.68) --
	(132.22,119.04) --
	(133.04,120.39) --
	(133.05,120.40) --
	(133.85,121.76) --
	(134.62,123.12) --
	(135.24,124.23) --
	(135.38,124.48) --
	(136.11,125.84) --
	(136.82,127.20) --
	(137.44,128.42) --
	(137.51,128.56) --
	(138.18,129.92) --
	(138.83,131.28) --
	(139.47,132.64) --
	(139.63,133.00) --
	(140.09,134.00) --
	(140.69,135.36) --
	(141.28,136.72) --
	(141.83,138.03) --
	(141.85,138.08) --
	(142.41,139.43) --
	(142.95,140.79) --
	(143.48,142.15) --
	(144.00,143.51) --
	(144.03,143.59) --
	(144.50,144.87) --
	(145.00,146.23) --
	(145.48,147.59) --
	(145.95,148.95) --
	(146.23,149.76) --
	(146.41,150.31) --
	(146.86,151.67) --
	(147.30,153.03) --
	(147.73,154.39) --
	(148.15,155.75) --
	(148.42,156.65) --
	(148.56,157.11) --
	(148.96,158.47) --
	(149.36,159.83) --
	(149.74,161.18) --
	(150.12,162.54) --
	(150.49,163.90) --
	(150.62,164.38) --
	(150.86,165.26) --
	(151.21,166.62) --
	(151.56,167.98) --
	(151.90,169.34) --
	(152.24,170.70) --
	(152.57,172.06) --
	(152.82,173.12) --
	(152.89,173.42) --
	(153.20,174.78) --
	(153.51,176.14) --
	(153.82,177.50) --
	(154.12,178.86) --
	(154.41,180.22) --
	(154.70,181.57) --
	(154.98,182.93) --
	(155.02,183.09) --
	(155.26,184.29) --
	(155.53,185.65) --
	(155.80,187.01) --
	(156.06,188.37) --
	(156.32,189.73) --
	(156.58,191.09) --
	(156.83,192.45) --
	(157.08,193.81) --
	(157.21,194.57) --
	(157.32,195.17) --
	(157.56,196.53) --
	(157.79,197.89) --
	(158.02,199.25) --
	(158.25,200.61) --
	(158.47,201.97) --
	(158.69,203.32) --
	(158.91,204.68) --
	(159.12,206.04) --
	(159.33,207.40) --
	(159.41,207.91) --
	(159.54,208.76) --
	(159.74,210.12) --
	(159.94,211.48) --
	(160.14,212.84) --
	(160.33,214.20) --
	(160.53,215.56) --
	(160.71,216.92) --
	(160.90,218.28) --
	(161.09,219.64) --
	(161.27,221.00) --
	(161.45,222.36) --
	(161.61,223.61) --
	(161.62,223.71) --
	(161.79,225.07) --
	(161.96,226.43) --
	(162.13,227.79) --
	(162.30,229.15) --
	(162.46,230.51) --
	(162.62,231.87) --
	(162.78,233.23) --
	(162.94,234.59) --
	(163.10,235.95) --
	(163.25,237.31) --
	(163.40,238.67) --
	(163.55,240.03) --
	(163.70,241.39) --
	(163.81,242.37) --
	(163.85,242.75) --
	(163.99,244.11) --
	(164.13,245.46) --
	(164.27,246.82) --
	(164.41,248.18) --
	(164.54,249.54) --
	(164.68,250.90) --
	(164.81,252.26) --
	(164.94,253.62) --
	(165.07,254.98) --
	(165.20,256.34) --
	(165.33,257.70) --
	(165.46,259.06) --
	(165.58,260.42) --
	(165.70,261.78) --
	(165.83,263.14) --
	(165.95,264.50) --
	(166.00,265.15) --
	(166.06,265.86) --
	(166.18,267.21) --
	(166.30,268.57) --
	(166.41,269.93) --
	(166.52,271.29) --
	(166.63,272.65) --
	(166.74,274.01) --
	(166.85,275.37) --
	(166.96,276.73) --
	(167.07,278.09) --
	(167.17,279.45) --
	(167.28,280.81) --
	(167.38,282.17) --
	(167.48,283.53) --
	(167.59,284.89) --
	(167.69,286.25) --
	(167.79,287.60) --
	(167.88,288.96) --
	(167.98,290.32) --
	(168.08,291.68) --
	(168.17,293.04) --
	(168.20,293.43) --
	(168.27,294.40) --
	(168.36,295.76) --
	(168.45,297.12) --
	(168.54,298.48) --
	(168.63,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41, 87.12) --
	(335.33, 87.78) --
	(335.21, 87.82) --
	(333.01, 88.53) --
	(331.22, 89.14) --
	(330.81, 89.28) --
	(328.62, 90.04) --
	(327.34, 90.50) --
	(326.42, 90.83) --
	(324.22, 91.65) --
	(323.68, 91.86) --
	(322.02, 92.50) --
	(320.22, 93.22) --
	(319.83, 93.38) --
	(317.63, 94.28) --
	(316.94, 94.58) --
	(315.43, 95.23) --
	(313.83, 95.94) --
	(313.23, 96.20) --
	(311.04, 97.22) --
	(310.87, 97.29) --
	(308.84, 98.27) --
	(308.05, 98.65) --
	(306.64, 99.36) --
	(305.37,100.01) --
	(304.44,100.49) --
	(302.80,101.37) --
	(302.25,101.67) --
	(300.35,102.73) --
	(300.05,102.90) --
	(298.01,104.09) --
	(297.85,104.18) --
	(295.76,105.45) --
	(295.65,105.52) --
	(293.61,106.81) --
	(293.46,106.91) --
	(291.54,108.17) --
	(291.26,108.36) --
	(289.55,109.53) --
	(289.06,109.87) --
	(287.64,110.89) --
	(286.86,111.46) --
	(285.81,112.25) --
	(284.67,113.12) --
	(284.04,113.61) --
	(282.47,114.85) --
	(282.33,114.97) --
	(280.68,116.33) --
	(280.27,116.67) --
	(279.10,117.68) --
	(278.07,118.58) --
	(277.56,119.04) --
	(276.08,120.40) --
	(275.88,120.59) --
	(274.64,121.76) --
	(273.68,122.70) --
	(273.26,123.12) --
	(271.91,124.48) --
	(271.48,124.93) --
	(270.61,125.84) --
	(269.35,127.20) --
	(269.28,127.27) --
	(268.13,128.56) --
	(267.09,129.75) --
	(266.94,129.92) --
	(265.79,131.28) --
	(264.89,132.37) --
	(264.67,132.64) --
	(263.58,134.00) --
	(262.69,135.14) --
	(262.52,135.36) --
	(261.50,136.72) --
	(260.50,138.08) --
	(260.49,138.08) --
	(259.52,139.43) --
	(258.58,140.79) --
	(258.30,141.20) --
	(257.65,142.15) --
	(256.75,143.51) --
	(256.10,144.53) --
	(255.88,144.87) --
	(255.02,146.23) --
	(254.19,147.59) --
	(253.90,148.07) --
	(253.38,148.95) --
	(252.58,150.31) --
	(251.81,151.67) --
	(251.70,151.86) --
	(251.05,153.03) --
	(250.31,154.39) --
	(249.59,155.75) --
	(249.51,155.91) --
	(248.89,157.11) --
	(248.20,158.47) --
	(247.52,159.83) --
	(247.31,160.26) --
	(246.86,161.18) --
	(246.22,162.54) --
	(245.59,163.90) --
	(245.11,164.94) --
	(244.97,165.26) --
	(244.36,166.62) --
	(243.77,167.98) --
	(243.19,169.34) --
	(242.92,169.99) --
	(242.62,170.70) --
	(242.06,172.06) --
	(241.52,173.42) --
	(240.98,174.78) --
	(240.72,175.46) --
	(240.46,176.14) --
	(239.94,177.50) --
	(239.43,178.86) --
	(238.94,180.22) --
	(238.52,181.39) --
	(238.45,181.57) --
	(237.97,182.93) --
	(237.51,184.29) --
	(237.05,185.65) --
	(236.59,187.01) --
	(236.32,187.84) --
	(236.15,188.37) --
	(235.71,189.73) --
	(235.29,191.09) --
	(234.86,192.45) --
	(234.45,193.81) --
	(234.13,194.90) --
	(234.05,195.17) --
	(233.64,196.53) --
	(233.25,197.89) --
	(232.86,199.25) --
	(232.49,200.61) --
	(232.11,201.97) --
	(231.93,202.65) --
	(231.74,203.32) --
	(231.38,204.68) --
	(231.02,206.04) --
	(230.67,207.40) --
	(230.33,208.76) --
	(229.99,210.12) --
	(229.73,211.18) --
	(229.66,211.48) --
	(229.33,212.84) --
	(229.00,214.20) --
	(228.68,215.56) --
	(228.37,216.92) --
	(228.06,218.28) --
	(227.75,219.64) --
	(227.53,220.64) --
	(227.45,221.00) --
	(227.16,222.36) --
	(226.86,223.71) --
	(226.58,225.07) --
	(226.29,226.43) --
	(226.01,227.79) --
	(225.74,229.15) --
	(225.47,230.51) --
	(225.34,231.17) --
	(225.20,231.87) --
	(224.93,233.23) --
	(224.67,234.59) --
	(224.41,235.95) --
	(224.16,237.31) --
	(223.91,238.67) --
	(223.66,240.03) --
	(223.42,241.39) --
	(223.18,242.75) --
	(223.14,242.97) --
	(222.94,244.11) --
	(222.70,245.46) --
	(222.47,246.82) --
	(222.24,248.18) --
	(222.02,249.54) --
	(221.79,250.90) --
	(221.57,252.26) --
	(221.36,253.62) --
	(221.14,254.98) --
	(220.94,256.29) --
	(220.93,256.34) --
	(220.72,257.70) --
	(220.51,259.06) --
	(220.31,260.42) --
	(220.11,261.78) --
	(219.91,263.14) --
	(219.71,264.50) --
	(219.51,265.86) --
	(219.32,267.21) --
	(219.13,268.57) --
	(218.94,269.93) --
	(218.76,271.29) --
	(218.74,271.42) --
	(218.58,272.65) --
	(218.39,274.01) --
	(218.21,275.37) --
	(218.03,276.73) --
	(217.86,278.09) --
	(217.68,279.45) --
	(217.51,280.81) --
	(217.34,282.17) --
	(217.17,283.53) --
	(217.01,284.89) --
	(216.84,286.25) --
	(216.68,287.60) --
	(216.55,288.77) --
	(216.52,288.96) --
	(216.36,290.32) --
	(216.20,291.68) --
	(216.05,293.04) --
	(215.89,294.40) --
	(215.74,295.76) --
	(215.59,297.12) --
	(215.44,298.48) --
	(215.29,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,152.71) --
	(336.93,153.03) --
	(335.21,154.21) --
	(334.95,154.39) --
	(333.03,155.75) --
	(333.01,155.76) --
	(331.16,157.11) --
	(330.81,157.36) --
	(329.34,158.47) --
	(328.62,159.01) --
	(327.57,159.83) --
	(326.42,160.73) --
	(325.84,161.18) --
	(324.22,162.50) --
	(324.16,162.54) --
	(322.53,163.90) --
	(322.02,164.33) --
	(320.93,165.26) --
	(319.83,166.22) --
	(319.38,166.62) --
	(317.86,167.98) --
	(317.63,168.19) --
	(316.38,169.34) --
	(315.43,170.22) --
	(314.93,170.70) --
	(313.52,172.06) --
	(313.23,172.34) --
	(312.14,173.42) --
	(311.04,174.53) --
	(310.79,174.78) --
	(309.47,176.14) --
	(308.84,176.80) --
	(308.18,177.50) --
	(306.92,178.86) --
	(306.64,179.16) --
	(305.69,180.22) --
	(304.48,181.57) --
	(304.44,181.61) --
	(303.30,182.93) --
	(302.25,184.17) --
	(302.14,184.29) --
	(301.01,185.65) --
	(300.05,186.82) --
	(299.90,187.01) --
	(298.81,188.37) --
	(297.85,189.59) --
	(297.74,189.73) --
	(296.69,191.09) --
	(295.67,192.45) --
	(295.65,192.47) --
	(294.66,193.81) --
	(293.68,195.17) --
	(293.46,195.48) --
	(292.71,196.53) --
	(291.76,197.89) --
	(291.26,198.61) --
	(290.83,199.25) --
	(289.91,200.61) --
	(289.06,201.89) --
	(289.01,201.97) --
	(288.13,203.32) --
	(287.26,204.68) --
	(286.86,205.32) --
	(286.41,206.04) --
	(285.57,207.40) --
	(284.75,208.76) --
	(284.67,208.90) --
	(283.94,210.12) --
	(283.15,211.48) --
	(282.47,212.66) --
	(282.36,212.84) --
	(281.59,214.20) --
	(280.84,215.56) --
	(280.27,216.59) --
	(280.10,216.92) --
	(279.36,218.28) --
	(278.64,219.64) --
	(278.07,220.73) --
	(277.93,221.00) --
	(277.23,222.36) --
	(276.55,223.71) --
	(275.88,225.07) --
	(275.87,225.07) --
	(275.21,226.43) --
	(274.55,227.79) --
	(273.90,229.15) --
	(273.68,229.63) --
	(273.27,230.51) --
	(272.64,231.87) --
	(272.02,233.23) --
	(271.48,234.44) --
	(271.42,234.59) --
	(270.81,235.95) --
	(270.22,237.31) --
	(269.64,238.67) --
	(269.28,239.51) --
	(269.07,240.03) --
	(268.50,241.39) --
	(267.94,242.75) --
	(267.39,244.11) --
	(267.09,244.86) --
	(266.85,245.46) --
	(266.31,246.82) --
	(265.78,248.18) --
	(265.26,249.54) --
	(264.89,250.52) --
	(264.75,250.90) --
	(264.24,252.26) --
	(263.73,253.62) --
	(263.24,254.98) --
	(262.76,256.34) --
	(262.69,256.52) --
	(262.27,257.70) --
	(261.79,259.06) --
	(261.33,260.42) --
	(260.86,261.78) --
	(260.49,262.87) --
	(260.41,263.14) --
	(259.95,264.50) --
	(259.50,265.86) --
	(259.06,267.21) --
	(258.63,268.57) --
	(258.30,269.63) --
	(258.20,269.93) --
	(257.77,271.29) --
	(257.35,272.65) --
	(256.94,274.01) --
	(256.53,275.37) --
	(256.12,276.73) --
	(256.10,276.81) --
	(255.72,278.09) --
	(255.32,279.45) --
	(254.93,280.81) --
	(254.55,282.17) --
	(254.16,283.53) --
	(253.90,284.47) --
	(253.79,284.89) --
	(253.41,286.25) --
	(253.04,287.60) --
	(252.67,288.96) --
	(252.31,290.32) --
	(251.96,291.68) --
	(251.70,292.66) --
	(251.61,293.04) --
	(251.25,294.40) --
	(250.91,295.76) --
	(250.56,297.12) --
	(250.23,298.48) --
	(249.89,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,201.89) --
	(337.33,201.97) --
	(335.91,203.32) --
	(335.21,204.00) --
	(334.51,204.68) --
	(333.14,206.04) --
	(333.01,206.18) --
	(331.80,207.40) --
	(330.81,208.42) --
	(330.49,208.76) --
	(329.20,210.12) --
	(328.62,210.74) --
	(327.93,211.48) --
	(326.69,212.84) --
	(326.42,213.15) --
	(325.48,214.20) --
	(324.28,215.56) --
	(324.22,215.63) --
	(323.11,216.92) --
	(322.02,218.20) --
	(321.96,218.28) --
	(320.83,219.64) --
	(319.83,220.86) --
	(319.72,221.00) --
	(318.62,222.36) --
	(317.63,223.62) --
	(317.55,223.71) --
	(316.50,225.07) --
	(315.46,226.43) --
	(315.43,226.47) --
	(314.44,227.79) --
	(313.44,229.15) --
	(313.23,229.43) --
	(312.45,230.51) --
	(311.48,231.87) --
	(311.04,232.51) --
	(310.53,233.23) --
	(309.59,234.59) --
	(308.84,235.70) --
	(308.67,235.95) --
	(307.76,237.31) --
	(306.86,238.67) --
	(306.64,239.01) --
	(305.98,240.03) --
	(305.11,241.39) --
	(304.44,242.45) --
	(304.26,242.75) --
	(303.42,244.11) --
	(302.59,245.46) --
	(302.25,246.03) --
	(301.77,246.82) --
	(300.97,248.18) --
	(300.18,249.54) --
	(300.05,249.76) --
	(299.39,250.90) --
	(298.62,252.26) --
	(297.86,253.62) --
	(297.85,253.64) --
	(297.11,254.98) --
	(296.37,256.34) --
	(295.65,257.69) --
	(295.65,257.70) --
	(294.93,259.06) --
	(294.22,260.42) --
	(293.52,261.78) --
	(293.46,261.90) --
	(292.83,263.14) --
	(292.15,264.50) --
	(291.48,265.86) --
	(291.26,266.31) --
	(290.82,267.21) --
	(290.16,268.57) --
	(289.52,269.93) --
	(289.06,270.90) --
	(288.88,271.29) --
	(288.25,272.65) --
	(287.63,274.01) --
	(287.02,275.37) --
	(286.86,275.71) --
	(286.41,276.73) --
	(285.81,278.09) --
	(285.22,279.45) --
	(284.67,280.74) --
	(284.64,280.81) --
	(284.06,282.17) --
	(283.49,283.53) --
	(282.93,284.89) --
	(282.47,286.01) --
	(282.37,286.25) --
	(281.82,287.60) --
	(281.28,288.96) --
	(280.74,290.32) --
	(280.27,291.53) --
	(280.21,291.68) --
	(279.69,293.04) --
	(279.17,294.40) --
	(278.66,295.76) --
	(278.15,297.12) --
	(278.07,297.33) --
	(277.65,298.48) --
	(277.15,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,240.15) --
	(336.34,241.39) --
	(335.21,242.73) --
	(335.19,242.75) --
	(334.06,244.11) --
	(333.01,245.39) --
	(332.95,245.46) --
	(331.85,246.82) --
	(330.81,248.14) --
	(330.78,248.18) --
	(329.72,249.54) --
	(328.67,250.90) --
	(328.62,250.98) --
	(327.65,252.26) --
	(326.64,253.62) --
	(326.42,253.92) --
	(325.64,254.98) --
	(324.66,256.34) --
	(324.22,256.95) --
	(323.69,257.70) --
	(322.74,259.06) --
	(322.02,260.10) --
	(321.80,260.42) --
	(320.88,261.78) --
	(319.97,263.14) --
	(319.83,263.35) --
	(319.07,264.50) --
	(318.19,265.86) --
	(317.63,266.73) --
	(317.32,267.21) --
	(316.45,268.57) --
	(315.61,269.93) --
	(315.43,270.22) --
	(314.77,271.29) --
	(313.95,272.65) --
	(313.23,273.85) --
	(313.14,274.01) --
	(312.33,275.37) --
	(311.54,276.73) --
	(311.04,277.60) --
	(310.76,278.09) --
	(309.98,279.45) --
	(309.23,280.81) --
	(308.84,281.51) --
	(308.47,282.17) --
	(307.73,283.53) --
	(307.00,284.89) --
	(306.64,285.56) --
	(306.28,286.25) --
	(305.56,287.60) --
	(304.86,288.96) --
	(304.44,289.77) --
	(304.16,290.32) --
	(303.47,291.68) --
	(302.79,293.04) --
	(302.25,294.16) --
	(302.13,294.40) --
	(301.46,295.76) --
	(300.80,297.12) --
	(300.16,298.48) --
	(300.05,298.71) --
	(299.52,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,270.76) --
	(337.00,271.29) --
	(335.98,272.65) --
	(335.21,273.71) --
	(334.99,274.01) --
	(334.00,275.37) --
	(333.03,276.73) --
	(333.01,276.76) --
	(332.07,278.09) --
	(331.13,279.45) --
	(330.81,279.91) --
	(330.20,280.81) --
	(329.28,282.17) --
	(328.62,283.17) --
	(328.38,283.53) --
	(327.48,284.89) --
	(326.60,286.25) --
	(326.42,286.53) --
	(325.73,287.60) --
	(324.87,288.96) --
	(324.22,290.02) --
	(324.03,290.32) --
	(323.19,291.68) --
	(322.37,293.04) --
	(322.02,293.62) --
	(321.55,294.40) --
	(320.75,295.76) --
	(319.96,297.12) --
	(319.83,297.35) --
	(319.17,298.48) --
	(318.40,299.84);

\path[draw=drawColor,line width= 0.6pt,line join=round] (337.41,295.80) --
	(336.50,297.12) --
	(335.59,298.48) --
	(335.21,299.06) --
	(334.69,299.84);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.16, 30.69) rectangle (338.79,302.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 54.84) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 82.03) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,109.22) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,136.40) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,163.59) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,190.78) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,217.97) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,245.15) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,272.34) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,299.53) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 30.69) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 57.87) --
	( 34.16, 57.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 85.06) --
	( 34.16, 85.06);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,112.25) --
	( 34.16,112.25);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,139.43) --
	( 34.16,139.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,166.62) --
	( 34.16,166.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,193.81) --
	( 34.16,193.81);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,221.00) --
	( 34.16,221.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,248.18) --
	( 34.16,248.18);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,275.37) --
	( 34.16,275.37);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,302.56) --
	( 34.16,302.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 27.94) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (186.47, 27.94) --
	(186.47, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (338.79, 27.94) --
	(338.79, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 34.16, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (186.47, 19.68) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (338.79, 19.68) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (186.47,  7.64) {Relative mean asymptomatic generation interval, $G_a/G_s$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,166.62) {Intrinsic proportion of asymptomatic transmission, $z$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (349.79, 25.48) rectangle (391.98,307.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at (355.29,  30.98) {
	\pgfimage[width= 14.45pt,height=256.07pt,interpolate=true]{figheatmap_R0_03_ras2}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24, 34.35) {1.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24, 88.41) {2.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,142.47) {3.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,196.53) {4.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (375.24,250.59) {5.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (355.29,293.62) {$\mathcal{R}_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29, 37.38) -- (358.18, 37.38);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29, 91.44) -- (358.18, 91.44);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,145.50) -- (358.18,145.50);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,199.56) -- (358.18,199.56);

\path[draw=drawColor,line width= 0.2pt,line join=round] (355.29,253.62) -- (358.18,253.62);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85, 37.38) -- (369.74, 37.38);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85, 91.44) -- (369.74, 91.44);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,145.50) -- (369.74,145.50);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,199.56) -- (369.74,199.56);

\path[draw=drawColor,line width= 0.2pt,line join=round] (366.85,253.62) -- (369.74,253.62);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (397.48,325.21);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.16,310.62) {B. Basic reproduction number};
\end{scope}
\end{tikzpicture}

\end{document}