### 2020-03-09 figheatmap_R0.tex
new figures!!!!!!

% Created by tikzDevice version 0.12.3 on 2020-03-09 13:46:06
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (361.35,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 34.16, 30.69) rectangle (302.65,266.42);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.16, 30.69) rectangle (302.65,266.42);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 42.47) --
	(302.65, 42.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 66.05) --
	(302.65, 66.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 89.62) --
	(302.65, 89.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,113.19) --
	(302.65,113.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,136.77) --
	(302.65,136.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,160.34) --
	(302.65,160.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,183.91) --
	(302.65,183.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,207.49) --
	(302.65,207.49);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,231.06) --
	(302.65,231.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,254.64) --
	(302.65,254.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (101.28, 30.69) --
	(101.28,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (235.53, 30.69) --
	(235.53,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	(302.65, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 54.26) --
	(302.65, 54.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 77.83) --
	(302.65, 77.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,101.41) --
	(302.65,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,124.98) --
	(302.65,124.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,148.55) --
	(302.65,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,172.13) --
	(302.65,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,195.70) --
	(302.65,195.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,219.28) --
	(302.65,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,242.85) --
	(302.65,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,266.42) --
	(302.65,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	( 34.16,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.41, 30.69) --
	(168.41,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.65, 30.69) --
	(302.65,266.42);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at ( 35.12,  31.86) {
	\pgfimage[width=267.28pt,height=233.38pt,interpolate=false]{figheatmap_R0_ras1}};
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,170.65) --
	( 35.66,172.13) --
	( 36.09,172.56) --
	( 37.99,174.49) --
	( 38.03,174.53) --
	( 39.97,176.56) --
	( 40.23,176.84) --
	( 41.90,178.66) --
	( 42.40,179.20) --
	( 43.84,180.82) --
	( 44.48,181.56) --
	( 45.78,183.06) --
	( 46.50,183.91) --
	( 47.71,185.37) --
	( 48.45,186.27) --
	( 49.65,187.76) --
	( 50.34,188.63) --
	( 51.59,190.23) --
	( 52.17,190.99) --
	( 53.52,192.79) --
	( 53.94,193.34) --
	( 55.46,195.44) --
	( 55.65,195.70) --
	( 57.31,198.06) --
	( 57.40,198.18) --
	( 58.92,200.42) --
	( 59.33,201.03) --
	( 60.49,202.77) --
	( 61.27,203.98) --
	( 62.01,205.13) --
	( 63.21,207.04) --
	( 63.48,207.49) --
	( 64.92,209.85) --
	( 65.14,210.22) --
	( 66.31,212.20) --
	( 67.08,213.53) --
	( 67.67,214.56) --
	( 68.99,216.92) --
	( 69.02,216.97) --
	( 70.27,219.28) --
	( 70.96,220.55) --
	( 71.52,221.63) --
	( 72.74,223.99) --
	( 72.89,224.28) --
	( 73.93,226.35) --
	( 74.83,228.17) --
	( 75.09,228.70) --
	( 76.22,231.06) --
	( 76.77,232.23) --
	( 77.32,233.42) --
	( 78.39,235.78) --
	( 78.70,236.46) --
	( 79.44,238.13) --
	( 80.47,240.49) --
	( 80.64,240.89) --
	( 81.47,242.85) --
	( 82.45,245.21) --
	( 82.58,245.52) --
	( 83.40,247.56) --
	( 84.34,249.92) --
	( 84.51,250.37) --
	( 85.25,252.28) --
	( 86.14,254.64) --
	( 86.45,255.45) --
	( 87.02,256.99) --
	( 87.87,259.35) --
	( 88.39,260.79) --
	( 88.71,261.71) --
	( 89.53,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 82.46) --
	( 34.41, 82.55) --
	( 36.09, 83.16) --
	( 38.03, 83.89) --
	( 39.97, 84.64) --
	( 40.63, 84.91) --
	( 41.90, 85.42) --
	( 43.84, 86.22) --
	( 45.78, 87.05) --
	( 46.27, 87.26) --
	( 47.71, 87.90) --
	( 49.65, 88.79) --
	( 51.42, 89.62) --
	( 51.59, 89.70) --
	( 53.52, 90.65) --
	( 55.46, 91.63) --
	( 56.14, 91.98) --
	( 57.40, 92.64) --
	( 59.33, 93.69) --
	( 60.48, 94.33) --
	( 61.27, 94.79) --
	( 63.21, 95.92) --
	( 64.49, 96.69) --
	( 65.14, 97.10) --
	( 67.08, 98.32) --
	( 68.20, 99.05) --
	( 69.02, 99.59) --
	( 70.96,100.92) --
	( 71.65,101.41) --
	( 72.89,102.30) --
	( 74.83,103.73) --
	( 74.87,103.76) --
	( 76.77,105.24) --
	( 77.87,106.12) --
	( 78.70,106.80) --
	( 80.64,108.44) --
	( 80.69,108.48) --
	( 82.58,110.15) --
	( 83.33,110.84) --
	( 84.51,111.95) --
	( 85.81,113.19) --
	( 86.45,113.83) --
	( 88.15,115.55) --
	( 88.39,115.80) --
	( 90.32,117.87) --
	( 90.36,117.91) --
	( 92.26,120.05) --
	( 92.45,120.27) --
	( 94.20,122.35) --
	( 94.42,122.62) --
	( 96.13,124.77) --
	( 96.30,124.98) --
	( 98.07,127.32) --
	( 98.08,127.34) --
	( 99.78,129.70) --
	(100.01,130.03) --
	(101.39,132.05) --
	(101.94,132.89) --
	(102.93,134.41) --
	(103.88,135.93) --
	(104.40,136.77) --
	(105.80,139.12) --
	(105.82,139.15) --
	(107.14,141.48) --
	(107.75,142.59) --
	(108.43,143.84) --
	(109.66,146.20) --
	(109.69,146.25) --
	(110.84,148.55) --
	(111.63,150.17) --
	(111.98,150.91) --
	(113.07,153.27) --
	(113.57,154.37) --
	(114.12,155.63) --
	(115.13,157.98) --
	(115.50,158.87) --
	(116.10,160.34) --
	(117.04,162.70) --
	(117.44,163.72) --
	(117.94,165.06) --
	(118.82,167.41) --
	(119.38,168.96) --
	(119.66,169.77) --
	(120.47,172.13) --
	(121.26,174.49) --
	(121.31,174.63) --
	(122.02,176.84) --
	(122.76,179.20) --
	(123.25,180.79) --
	(123.48,181.56) --
	(124.17,183.91) --
	(124.84,186.27) --
	(125.19,187.51) --
	(125.49,188.63) --
	(126.12,190.99) --
	(126.74,193.34) --
	(127.12,194.86) --
	(127.33,195.70) --
	(127.91,198.06) --
	(128.47,200.42) --
	(129.02,202.77) --
	(129.06,202.94) --
	(129.55,205.13) --
	(130.07,207.49) --
	(130.57,209.85) --
	(131.00,211.86) --
	(131.07,212.20) --
	(131.54,214.56) --
	(132.01,216.92) --
	(132.46,219.28) --
	(132.91,221.63) --
	(132.93,221.76) --
	(133.34,223.99) --
	(133.76,226.35) --
	(134.17,228.70) --
	(134.58,231.06) --
	(134.87,232.80) --
	(134.97,233.42) --
	(135.35,235.78) --
	(135.73,238.13) --
	(136.09,240.49) --
	(136.45,242.85) --
	(136.81,245.21) --
	(136.81,245.21) --
	(137.15,247.56) --
	(137.48,249.92) --
	(137.81,252.28) --
	(138.13,254.64) --
	(138.45,256.99) --
	(138.74,259.24) --
	(138.76,259.35) --
	(139.06,261.71) --
	(139.35,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44, 46.16) --
	(299.50, 46.36) --
	(297.56, 46.56) --
	(295.62, 46.77) --
	(293.69, 46.99) --
	(291.99, 47.19) --
	(291.75, 47.22) --
	(289.81, 47.45) --
	(287.88, 47.69) --
	(285.94, 47.93) --
	(284.00, 48.19) --
	(282.07, 48.46) --
	(280.13, 48.73) --
	(278.19, 49.02) --
	(276.26, 49.32) --
	(274.82, 49.54) --
	(274.32, 49.63) --
	(272.38, 49.94) --
	(270.45, 50.28) --
	(268.51, 50.62) --
	(266.57, 50.98) --
	(264.64, 51.36) --
	(262.70, 51.75) --
	(261.97, 51.90) --
	(260.76, 52.16) --
	(258.83, 52.58) --
	(256.89, 53.03) --
	(254.95, 53.49) --
	(253.02, 53.98) --
	(251.96, 54.26) --
	(251.08, 54.49) --
	(249.14, 55.03) --
	(247.20, 55.59) --
	(245.27, 56.19) --
	(243.92, 56.62) --
	(243.33, 56.81) --
	(241.39, 57.47) --
	(239.46, 58.16) --
	(237.52, 58.89) --
	(237.32, 58.97) --
	(235.58, 59.67) --
	(233.65, 60.49) --
	(231.79, 61.33) --
	(231.71, 61.37) --
	(229.77, 62.30) --
	(227.84, 63.29) --
	(227.09, 63.69) --
	(225.90, 64.35) --
	(223.96, 65.48) --
	(223.05, 66.05) --
	(222.03, 66.70) --
	(220.09, 68.01) --
	(219.54, 68.40) --
	(218.15, 69.42) --
	(216.45, 70.76) --
	(216.22, 70.95) --
	(214.28, 72.61) --
	(213.71, 73.12) --
	(212.34, 74.41) --
	(211.28, 75.48) --
	(210.41, 76.38) --
	(209.09, 77.83) --
	(208.47, 78.55) --
	(207.11, 80.19) --
	(206.53, 80.93) --
	(205.32, 82.55) --
	(204.59, 83.57) --
	(203.69, 84.91) --
	(202.66, 86.52) --
	(202.20, 87.26) --
	(200.83, 89.62) --
	(200.72, 89.81) --
	(199.56, 91.98) --
	(198.78, 93.53) --
	(198.39, 94.33) --
	(197.31, 96.69) --
	(196.85, 97.75) --
	(196.30, 99.05) --
	(195.36,101.41) --
	(194.91,102.60) --
	(194.48,103.76) --
	(193.66,106.12) --
	(192.97,108.21) --
	(192.89,108.48) --
	(192.16,110.84) --
	(191.48,113.19) --
	(191.04,114.78) --
	(190.83,115.55) --
	(190.22,117.91) --
	(189.64,120.27) --
	(189.10,122.58) --
	(189.09,122.62) --
	(188.57,124.98) --
	(188.07,127.34) --
	(187.60,129.70) --
	(187.16,132.00) --
	(187.15,132.05) --
	(186.72,134.41) --
	(186.31,136.77) --
	(185.92,139.12) --
	(185.55,141.48) --
	(185.23,143.59) --
	(185.19,143.84) --
	(184.84,146.20) --
	(184.51,148.55) --
	(184.20,150.91) --
	(183.89,153.27) --
	(183.60,155.63) --
	(183.32,157.98) --
	(183.29,158.20) --
	(183.04,160.34) --
	(182.78,162.70) --
	(182.53,165.06) --
	(182.28,167.41) --
	(182.04,169.77) --
	(181.82,172.13) --
	(181.60,174.49) --
	(181.38,176.84) --
	(181.35,177.19) --
	(181.18,179.20) --
	(180.98,181.56) --
	(180.78,183.91) --
	(180.59,186.27) --
	(180.41,188.63) --
	(180.23,190.99) --
	(180.06,193.34) --
	(179.89,195.70) --
	(179.73,198.06) --
	(179.57,200.42) --
	(179.42,202.77) --
	(179.42,202.85) --
	(179.27,205.13) --
	(179.13,207.49) --
	(178.98,209.85) --
	(178.85,212.20) --
	(178.71,214.56) --
	(178.58,216.92) --
	(178.45,219.28) --
	(178.33,221.63) --
	(178.21,223.99) --
	(178.09,226.35) --
	(177.97,228.70) --
	(177.86,231.06) --
	(177.75,233.42) --
	(177.64,235.78) --
	(177.54,238.13) --
	(177.48,239.48) --
	(177.44,240.49) --
	(177.33,242.85) --
	(177.24,245.21) --
	(177.14,247.56) --
	(177.04,249.92) --
	(176.95,252.28) --
	(176.86,254.64) --
	(176.77,256.99) --
	(176.69,259.35) --
	(176.60,261.71) --
	(176.52,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,101.56) --
	(299.50,102.46) --
	(297.56,103.39) --
	(296.81,103.76) --
	(295.62,104.35) --
	(293.69,105.35) --
	(292.23,106.12) --
	(291.75,106.38) --
	(289.81,107.44) --
	(287.98,108.48) --
	(287.88,108.54) --
	(285.94,109.67) --
	(284.03,110.84) --
	(284.00,110.85) --
	(282.07,112.07) --
	(280.34,113.19) --
	(280.13,113.34) --
	(278.19,114.64) --
	(276.90,115.55) --
	(276.26,116.00) --
	(274.32,117.41) --
	(273.66,117.91) --
	(272.38,118.88) --
	(270.62,120.27) --
	(270.45,120.40) --
	(268.51,121.99) --
	(267.76,122.62) --
	(266.57,123.64) --
	(265.05,124.98) --
	(264.64,125.36) --
	(262.70,127.15) --
	(262.50,127.34) --
	(260.76,129.02) --
	(260.08,129.70) --
	(258.83,130.97) --
	(257.79,132.05) --
	(256.89,133.01) --
	(255.61,134.41) --
	(254.95,135.14) --
	(253.53,136.77) --
	(253.02,137.37) --
	(251.56,139.12) --
	(251.08,139.71) --
	(249.68,141.48) --
	(249.14,142.17) --
	(247.88,143.84) --
	(247.20,144.75) --
	(246.16,146.20) --
	(245.27,147.46) --
	(244.52,148.55) --
	(243.33,150.32) --
	(242.94,150.91) --
	(241.43,153.27) --
	(241.39,153.33) --
	(239.98,155.63) --
	(239.46,156.51) --
	(238.59,157.98) --
	(237.52,159.86) --
	(237.25,160.34) --
	(235.97,162.70) --
	(235.58,163.42) --
	(234.73,165.06) --
	(233.65,167.19) --
	(233.54,167.41) --
	(232.38,169.77) --
	(231.71,171.19) --
	(231.27,172.13) --
	(230.20,174.49) --
	(229.77,175.45) --
	(229.17,176.84) --
	(228.17,179.20) --
	(227.84,180.00) --
	(227.20,181.56) --
	(226.26,183.91) --
	(225.90,184.85) --
	(225.35,186.27) --
	(224.47,188.63) --
	(223.96,190.04) --
	(223.62,190.99) --
	(222.80,193.34) --
	(222.03,195.61) --
	(222.00,195.70) --
	(221.22,198.06) --
	(220.46,200.42) --
	(220.09,201.61) --
	(219.73,202.77) --
	(219.01,205.13) --
	(218.32,207.49) --
	(218.15,208.07) --
	(217.65,209.85) --
	(216.99,212.20) --
	(216.35,214.56) --
	(216.22,215.07) --
	(215.73,216.92) --
	(215.12,219.28) --
	(214.53,221.63) --
	(214.28,222.66) --
	(213.95,223.99) --
	(213.39,226.35) --
	(212.84,228.70) --
	(212.34,230.92) --
	(212.31,231.06) --
	(211.79,233.42) --
	(211.28,235.78) --
	(210.78,238.13) --
	(210.41,239.95) --
	(210.29,240.49) --
	(209.82,242.85) --
	(209.35,245.21) --
	(208.90,247.56) --
	(208.47,249.86) --
	(208.46,249.92) --
	(208.02,252.28) --
	(207.59,254.64) --
	(207.18,256.99) --
	(206.77,259.35) --
	(206.53,260.78) --
	(206.38,261.71) --
	(205.98,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,145.87) --
	(301.00,146.20) --
	(299.50,147.34) --
	(297.94,148.55) --
	(297.56,148.86) --
	(295.62,150.42) --
	(295.03,150.91) --
	(293.69,152.04) --
	(292.26,153.27) --
	(291.75,153.71) --
	(289.81,155.44) --
	(289.60,155.63) --
	(287.88,157.22) --
	(287.07,157.98) --
	(285.94,159.07) --
	(284.64,160.34) --
	(284.00,160.98) --
	(282.32,162.70) --
	(282.07,162.96) --
	(280.13,165.02) --
	(280.09,165.06) --
	(278.19,167.14) --
	(277.95,167.41) --
	(276.26,169.35) --
	(275.90,169.77) --
	(274.32,171.65) --
	(273.92,172.13) --
	(272.38,174.03) --
	(272.02,174.49) --
	(270.45,176.51) --
	(270.19,176.84) --
	(268.51,179.08) --
	(268.42,179.20) --
	(266.72,181.56) --
	(266.57,181.76) --
	(265.07,183.91) --
	(264.64,184.55) --
	(263.48,186.27) --
	(262.70,187.46) --
	(261.95,188.63) --
	(260.76,190.50) --
	(260.46,190.99) --
	(259.02,193.34) --
	(258.83,193.68) --
	(257.63,195.70) --
	(256.89,196.99) --
	(256.28,198.06) --
	(254.97,200.42) --
	(254.95,200.46) --
	(253.70,202.77) --
	(253.02,204.09) --
	(252.47,205.13) --
	(251.28,207.49) --
	(251.08,207.89) --
	(250.12,209.85) --
	(249.14,211.88) --
	(248.99,212.20) --
	(247.89,214.56) --
	(247.20,216.07) --
	(246.83,216.92) --
	(245.79,219.28) --
	(245.27,220.48) --
	(244.78,221.63) --
	(243.79,223.99) --
	(243.33,225.12) --
	(242.84,226.35) --
	(241.90,228.70) --
	(241.39,230.02) --
	(240.99,231.06) --
	(240.10,233.42) --
	(239.46,235.18) --
	(239.24,235.78) --
	(238.39,238.13) --
	(237.57,240.49) --
	(237.52,240.64) --
	(236.76,242.85) --
	(235.98,245.21) --
	(235.58,246.42) --
	(235.21,247.56) --
	(234.46,249.92) --
	(233.73,252.28) --
	(233.65,252.55) --
	(233.01,254.64) --
	(232.31,256.99) --
	(231.71,259.05) --
	(231.62,259.35) --
	(230.95,261.71) --
	(230.29,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,182.13) --
	(299.64,183.91) --
	(299.50,184.06) --
	(297.56,186.05) --
	(297.35,186.27) --
	(295.62,188.11) --
	(295.15,188.63) --
	(293.69,190.24) --
	(293.02,190.99) --
	(291.75,192.43) --
	(290.97,193.34) --
	(289.81,194.70) --
	(288.98,195.70) --
	(287.88,197.05) --
	(287.07,198.06) --
	(285.94,199.48) --
	(285.21,200.42) --
	(284.00,201.99) --
	(283.42,202.77) --
	(282.07,204.60) --
	(281.68,205.13) --
	(280.13,207.30) --
	(280.00,207.49) --
	(278.37,209.85) --
	(278.19,210.10) --
	(276.78,212.20) --
	(276.26,213.00) --
	(275.25,214.56) --
	(274.32,216.02) --
	(273.76,216.92) --
	(272.38,219.15) --
	(272.31,219.28) --
	(270.90,221.63) --
	(270.45,222.41) --
	(269.53,223.99) --
	(268.51,225.79) --
	(268.20,226.35) --
	(266.90,228.70) --
	(266.57,229.32) --
	(265.64,231.06) --
	(264.64,232.99) --
	(264.41,233.42) --
	(263.22,235.78) --
	(262.70,236.82) --
	(262.05,238.13) --
	(260.91,240.49) --
	(260.76,240.81) --
	(259.80,242.85) --
	(258.83,244.98) --
	(258.72,245.21) --
	(257.66,247.56) --
	(256.89,249.34) --
	(256.64,249.92) --
	(255.63,252.28) --
	(254.95,253.90) --
	(254.65,254.64) --
	(253.68,256.99) --
	(253.02,258.67) --
	(252.75,259.35) --
	(251.83,261.71) --
	(251.08,263.67) --
	(250.93,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,212.35) --
	(299.58,214.56) --
	(299.50,214.66) --
	(297.67,216.92) --
	(297.56,217.05) --
	(295.82,219.28) --
	(295.62,219.52) --
	(294.02,221.63) --
	(293.69,222.07) --
	(292.27,223.99) --
	(291.75,224.70) --
	(290.57,226.35) --
	(289.81,227.43) --
	(288.93,228.70) --
	(287.88,230.24) --
	(287.33,231.06) --
	(285.94,233.16) --
	(285.77,233.42) --
	(284.25,235.78) --
	(284.00,236.17) --
	(282.78,238.13) --
	(282.07,239.30) --
	(281.34,240.49) --
	(280.13,242.54) --
	(279.95,242.85) --
	(278.58,245.21) --
	(278.19,245.89) --
	(277.25,247.56) --
	(276.26,249.38) --
	(275.96,249.92) --
	(274.70,252.28) --
	(274.32,252.99) --
	(273.46,254.64) --
	(272.38,256.75) --
	(272.26,256.99) --
	(271.08,259.35) --
	(270.45,260.66) --
	(269.94,261.71) --
	(268.81,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,237.91) --
	(301.27,238.13) --
	(299.54,240.49) --
	(299.50,240.56) --
	(297.86,242.85) --
	(297.56,243.28) --
	(296.23,245.21) --
	(295.62,246.10) --
	(294.64,247.56) --
	(293.69,249.00) --
	(293.09,249.92) --
	(291.75,252.01) --
	(291.58,252.28) --
	(290.11,254.64) --
	(289.81,255.12) --
	(288.67,256.99) --
	(287.88,258.33) --
	(287.27,259.35) --
	(285.94,261.65) --
	(285.91,261.71) --
	(284.57,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,259.83) --
	(300.18,261.71) --
	(299.50,262.75) --
	(298.64,264.07);

\path[draw=drawColor,line width= 0.9pt,dash pattern=on 4pt off 4pt ,line join=round] (168.41, 30.69) -- (168.41,266.42);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (164.30, 56.62) {$\mathcal{R}_0=2.5$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (235.53, 33.96) {underestimated};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (101.28, 33.96) {overestimated};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 52.62,148.55) circle (  4.64);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 59.53,135.32) {$\mathcal{R}_0=2.0$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (282.25,155.77) --
	(288.50,144.95) --
	(276.00,144.95) --
	cycle;

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (280.08,135.32) {$\mathcal{R}_0=3.1$};
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.16, 30.69) rectangle (302.65,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 51.23) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 74.80) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 98.38) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,121.95) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,145.52) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,169.10) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,192.67) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,216.25) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,239.82) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,263.39) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 30.69) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 54.26) --
	( 34.16, 54.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 77.83) --
	( 34.16, 77.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,101.41) --
	( 34.16,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,124.98) --
	( 34.16,124.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,148.55) --
	( 34.16,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,172.13) --
	( 34.16,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,195.70) --
	( 34.16,195.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,219.28) --
	( 34.16,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,242.85) --
	( 34.16,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,266.42) --
	( 34.16,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 27.94) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.41, 27.94) --
	(168.41, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.65, 27.94) --
	(302.65, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 34.16, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (168.41, 19.68) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.65, 19.68) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (168.41,  7.64) {Relative mean asymptomatic generation interval, $G_a/G_s$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,148.55) {Intrinsic proportion of asymptomatic transmission, $z$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (313.65, 28.75) rectangle (355.85,268.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at (319.15,  34.25) {
	\pgfimage[width= 14.45pt,height=213.40pt,interpolate=true]{figheatmap_R0_ras2}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11, 36.55) {1.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11, 81.60) {2.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,126.65) {3.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,171.70) {4.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,216.75) {5.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (319.15,254.21) {$\mathcal{R}_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15, 39.58) -- (322.05, 39.58);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15, 84.63) -- (322.05, 84.63);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,129.68) -- (322.05,129.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,174.73) -- (322.05,174.73);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,219.78) -- (322.05,219.78);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72, 39.58) -- (333.61, 39.58);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72, 84.63) -- (333.61, 84.63);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,129.68) -- (333.61,129.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,174.73) -- (333.61,174.73);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,219.78) -- (333.61,219.78);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.16,274.49) {B. Basic reproduction number};
\end{scope}
\end{tikzpicture}

\end{document}