### 2020-03-09 figheatmap_R0_03.tex

% Created by tikzDevice version 0.12.3 on 2020-03-09 13:22:38
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,289.08);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (361.35,289.08);
\end{scope}
\begin{scope}
\path[clip] ( 34.16, 30.69) rectangle (302.65,266.42);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 34.16, 30.69) rectangle (302.65,266.42);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 42.47) --
	(302.65, 42.47);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 66.05) --
	(302.65, 66.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16, 89.62) --
	(302.65, 89.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,113.19) --
	(302.65,113.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,136.77) --
	(302.65,136.77);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,160.34) --
	(302.65,160.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,183.91) --
	(302.65,183.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,207.49) --
	(302.65,207.49);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,231.06) --
	(302.65,231.06);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 34.16,254.64) --
	(302.65,254.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (101.28, 30.69) --
	(101.28,266.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] (235.53, 30.69) --
	(235.53,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	(302.65, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 54.26) --
	(302.65, 54.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 77.83) --
	(302.65, 77.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,101.41) --
	(302.65,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,124.98) --
	(302.65,124.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,148.55) --
	(302.65,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,172.13) --
	(302.65,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,195.70) --
	(302.65,195.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,219.28) --
	(302.65,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,242.85) --
	(302.65,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,266.42) --
	(302.65,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 30.69) --
	( 34.16,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.41, 30.69) --
	(168.41,266.42);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.65, 30.69) --
	(302.65,266.42);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at ( 35.12,  32.45) {
	\pgfimage[width=267.28pt,height=232.20pt,interpolate=false]{figheatmap_R0_03_ras1}};
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16,167.98) --
	( 34.80,168.59) --
	( 36.03,169.77) --
	( 36.09,169.83) --
	( 37.22,170.95) --
	( 38.03,171.75) --
	( 38.40,172.13) --
	( 39.56,173.31) --
	( 39.97,173.73) --
	( 40.69,174.49) --
	( 41.80,175.66) --
	( 41.90,175.77) --
	( 42.89,176.84) --
	( 43.84,177.88) --
	( 43.97,178.02) --
	( 45.02,179.20) --
	( 45.78,180.06) --
	( 46.06,180.38) --
	( 47.07,181.56) --
	( 47.71,182.31) --
	( 48.07,182.74) --
	( 49.06,183.91) --
	( 49.65,184.64) --
	( 50.02,185.09) --
	( 50.97,186.27) --
	( 51.59,187.04) --
	( 51.91,187.45) --
	( 52.83,188.63) --
	( 53.52,189.54) --
	( 53.73,189.81) --
	( 54.62,190.99) --
	( 55.46,192.12) --
	( 55.50,192.17) --
	( 56.36,193.34) --
	( 57.21,194.52) --
	( 57.40,194.79) --
	( 58.04,195.70) --
	( 58.86,196.88) --
	( 59.33,197.56) --
	( 59.67,198.06) --
	( 60.47,199.24) --
	( 61.26,200.42) --
	( 61.27,200.44) --
	( 62.03,201.60) --
	( 62.79,202.77) --
	( 63.21,203.43) --
	( 63.54,203.95) --
	( 64.28,205.13) --
	( 65.01,206.31) --
	( 65.14,206.53) --
	( 65.73,207.49) --
	( 66.43,208.67) --
	( 67.08,209.75) --
	( 67.14,209.85) --
	( 67.82,211.02) --
	( 68.50,212.20) --
	( 69.02,213.11) --
	( 69.17,213.38) --
	( 69.83,214.56) --
	( 70.48,215.74) --
	( 70.96,216.60) --
	( 71.13,216.92) --
	( 71.76,218.10) --
	( 72.38,219.28) --
	( 72.89,220.24) --
	( 73.00,220.45) --
	( 73.61,221.63) --
	( 74.21,222.81) --
	( 74.81,223.99) --
	( 74.83,224.03) --
	( 75.39,225.17) --
	( 75.97,226.35) --
	( 76.54,227.53) --
	( 76.77,227.99) --
	( 77.11,228.70) --
	( 77.66,229.88) --
	( 78.21,231.06) --
	( 78.70,232.12) --
	( 78.76,232.24) --
	( 79.29,233.42) --
	( 79.82,234.60) --
	( 80.35,235.78) --
	( 80.64,236.44) --
	( 80.86,236.96) --
	( 81.37,238.13) --
	( 81.88,239.31) --
	( 82.38,240.49) --
	( 82.58,240.96) --
	( 82.87,241.67) --
	( 83.36,242.85) --
	( 83.84,244.03) --
	( 84.32,245.21) --
	( 84.51,245.69) --
	( 84.79,246.39) --
	( 85.25,247.56) --
	( 85.71,248.74) --
	( 86.17,249.92) --
	( 86.45,250.65) --
	( 86.62,251.10) --
	( 87.06,252.28) --
	( 87.50,253.46) --
	( 87.94,254.64) --
	( 88.37,255.81) --
	( 88.39,255.86) --
	( 88.79,256.99) --
	( 89.21,258.17) --
	( 89.63,259.35) --
	( 90.04,260.53) --
	( 90.32,261.33) --
	( 90.45,261.71) --
	( 90.85,262.89) --
	( 91.25,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 58.75) --
	( 35.33, 58.97) --
	( 36.09, 59.12) --
	( 38.03, 59.52) --
	( 39.97, 59.92) --
	( 41.05, 60.15) --
	( 41.90, 60.34) --
	( 43.84, 60.77) --
	( 45.78, 61.21) --
	( 46.28, 61.33) --
	( 47.71, 61.67) --
	( 49.65, 62.15) --
	( 51.08, 62.51) --
	( 51.59, 62.64) --
	( 53.52, 63.15) --
	( 55.46, 63.68) --
	( 55.50, 63.69) --
	( 57.40, 64.23) --
	( 59.33, 64.79) --
	( 59.59, 64.87) --
	( 61.27, 65.38) --
	( 63.21, 65.99) --
	( 63.38, 66.05) --
	( 65.14, 66.63) --
	( 66.91, 67.23) --
	( 67.08, 67.28) --
	( 69.02, 67.97) --
	( 70.21, 68.40) --
	( 70.96, 68.68) --
	( 72.89, 69.43) --
	( 73.29, 69.58) --
	( 74.83, 70.20) --
	( 76.18, 70.76) --
	( 76.77, 71.01) --
	( 78.70, 71.85) --
	( 78.89, 71.94) --
	( 80.64, 72.74) --
	( 81.45, 73.12) --
	( 82.58, 73.66) --
	( 83.86, 74.30) --
	( 84.51, 74.63) --
	( 86.14, 75.48) --
	( 86.45, 75.64) --
	( 88.29, 76.65) --
	( 88.39, 76.71) --
	( 90.32, 77.82) --
	( 90.34, 77.83) --
	( 92.26, 79.00) --
	( 92.28, 79.01) --
	( 94.12, 80.19) --
	( 94.20, 80.24) --
	( 95.88, 81.37) --
	( 96.13, 81.55) --
	( 97.55, 82.55) --
	( 98.07, 82.92) --
	( 99.15, 83.73) --
	(100.01, 84.38) --
	(100.67, 84.91) --
	(101.94, 85.93) --
	(102.13, 86.08) --
	(103.53, 87.26) --
	(103.88, 87.57) --
	(104.87, 88.44) --
	(105.82, 89.31) --
	(106.15, 89.62) --
	(107.38, 90.80) --
	(107.75, 91.16) --
	(108.57, 91.98) --
	(109.69, 93.14) --
	(109.70, 93.16) --
	(110.80, 94.33) --
	(111.63, 95.26) --
	(111.85, 95.51) --
	(112.87, 96.69) --
	(113.57, 97.52) --
	(113.85, 97.87) --
	(114.80, 99.05) --
	(115.50, 99.96) --
	(115.71,100.23) --
	(116.59,101.41) --
	(117.44,102.58) --
	(117.44,102.59) --
	(118.27,103.76) --
	(119.07,104.94) --
	(119.38,105.41) --
	(119.84,106.12) --
	(120.59,107.30) --
	(121.31,108.47) --
	(121.32,108.48) --
	(122.02,109.66) --
	(122.70,110.84) --
	(123.25,111.80) --
	(123.37,112.02) --
	(124.01,113.19) --
	(124.64,114.37) --
	(125.19,115.43) --
	(125.25,115.55) --
	(125.84,116.73) --
	(126.41,117.91) --
	(126.98,119.09) --
	(127.12,119.40) --
	(127.52,120.27) --
	(128.05,121.44) --
	(128.57,122.62) --
	(129.06,123.77) --
	(129.07,123.80) --
	(129.57,124.98) --
	(130.04,126.16) --
	(130.51,127.34) --
	(130.97,128.52) --
	(131.00,128.59) --
	(131.41,129.70) --
	(131.85,130.87) --
	(132.27,132.05) --
	(132.69,133.23) --
	(132.93,133.93) --
	(133.10,134.41) --
	(133.49,135.59) --
	(133.88,136.77) --
	(134.26,137.95) --
	(134.63,139.12) --
	(134.87,139.90) --
	(134.99,140.30) --
	(135.35,141.48) --
	(135.69,142.66) --
	(136.03,143.84) --
	(136.37,145.02) --
	(136.69,146.20) --
	(136.81,146.61) --
	(137.01,147.38) --
	(137.33,148.55) --
	(137.63,149.73) --
	(137.93,150.91) --
	(138.23,152.09) --
	(138.52,153.27) --
	(138.74,154.19) --
	(138.80,154.45) --
	(139.08,155.63) --
	(139.36,156.81) --
	(139.62,157.98) --
	(139.89,159.16) --
	(140.15,160.34) --
	(140.40,161.52) --
	(140.65,162.70) --
	(140.68,162.84) --
	(140.90,163.88) --
	(141.14,165.06) --
	(141.37,166.23) --
	(141.60,167.41) --
	(141.83,168.59) --
	(142.06,169.77) --
	(142.28,170.95) --
	(142.50,172.13) --
	(142.62,172.78) --
	(142.71,173.31) --
	(142.92,174.49) --
	(143.13,175.66) --
	(143.33,176.84) --
	(143.53,178.02) --
	(143.73,179.20) --
	(143.92,180.38) --
	(144.11,181.56) --
	(144.30,182.74) --
	(144.49,183.91) --
	(144.55,184.35) --
	(144.67,185.09) --
	(144.85,186.27) --
	(145.02,187.45) --
	(145.20,188.63) --
	(145.37,189.81) --
	(145.54,190.99) --
	(145.70,192.17) --
	(145.87,193.34) --
	(146.03,194.52) --
	(146.19,195.70) --
	(146.35,196.88) --
	(146.49,197.97) --
	(146.50,198.06) --
	(146.65,199.24) --
	(146.80,200.42) --
	(146.95,201.60) --
	(147.10,202.77) --
	(147.24,203.95) --
	(147.39,205.13) --
	(147.53,206.31) --
	(147.66,207.49) --
	(147.80,208.67) --
	(147.94,209.85) --
	(148.07,211.02) --
	(148.20,212.20) --
	(148.33,213.38) --
	(148.43,214.23) --
	(148.46,214.56) --
	(148.59,215.74) --
	(148.71,216.92) --
	(148.84,218.10) --
	(148.96,219.28) --
	(149.08,220.45) --
	(149.20,221.63) --
	(149.31,222.81) --
	(149.43,223.99) --
	(149.55,225.17) --
	(149.66,226.35) --
	(149.77,227.53) --
	(149.88,228.70) --
	(149.99,229.88) --
	(150.10,231.06) --
	(150.21,232.24) --
	(150.31,233.42) --
	(150.36,233.99) --
	(150.42,234.60) --
	(150.52,235.78) --
	(150.62,236.96) --
	(150.72,238.13) --
	(150.82,239.31) --
	(150.92,240.49) --
	(151.02,241.67) --
	(151.11,242.85) --
	(151.21,244.03) --
	(151.30,245.21) --
	(151.39,246.39) --
	(151.49,247.56) --
	(151.58,248.74) --
	(151.67,249.92) --
	(151.76,251.10) --
	(151.85,252.28) --
	(151.93,253.46) --
	(152.02,254.64) --
	(152.11,255.81) --
	(152.19,256.99) --
	(152.28,258.17) --
	(152.30,258.51) --
	(152.36,259.35) --
	(152.44,260.53) --
	(152.52,261.71) --
	(152.60,262.89) --
	(152.68,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44, 79.62) --
	(299.61, 80.19) --
	(299.50, 80.22) --
	(297.56, 80.85) --
	(295.98, 81.37) --
	(295.62, 81.49) --
	(293.69, 82.15) --
	(292.56, 82.55) --
	(291.75, 82.84) --
	(289.81, 83.55) --
	(289.34, 83.73) --
	(287.88, 84.28) --
	(286.29, 84.91) --
	(285.94, 85.04) --
	(284.00, 85.83) --
	(283.40, 86.08) --
	(282.07, 86.65) --
	(280.65, 87.26) --
	(280.13, 87.49) --
	(278.19, 88.37) --
	(278.05, 88.44) --
	(276.26, 89.28) --
	(275.56, 89.62) --
	(274.32, 90.23) --
	(273.20, 90.80) --
	(272.38, 91.22) --
	(270.94, 91.98) --
	(270.45, 92.24) --
	(268.78, 93.16) --
	(268.51, 93.30) --
	(266.71, 94.33) --
	(266.57, 94.41) --
	(264.73, 95.51) --
	(264.64, 95.57) --
	(262.83, 96.69) --
	(262.70, 96.78) --
	(261.01, 97.87) --
	(260.76, 98.03) --
	(259.26, 99.05) --
	(258.83, 99.35) --
	(257.57,100.23) --
	(256.89,100.72) --
	(255.96,101.41) --
	(254.95,102.16) --
	(254.40,102.59) --
	(253.02,103.67) --
	(252.89,103.76) --
	(251.44,104.94) --
	(251.08,105.24) --
	(250.04,106.12) --
	(249.14,106.90) --
	(248.69,107.30) --
	(247.38,108.48) --
	(247.20,108.64) --
	(246.12,109.66) --
	(245.27,110.47) --
	(244.90,110.84) --
	(243.71,112.02) --
	(243.33,112.40) --
	(242.56,113.19) --
	(241.45,114.37) --
	(241.39,114.44) --
	(240.37,115.55) --
	(239.46,116.58) --
	(239.33,116.73) --
	(238.31,117.91) --
	(237.52,118.85) --
	(237.33,119.09) --
	(236.37,120.27) --
	(235.58,121.26) --
	(235.44,121.44) --
	(234.53,122.62) --
	(233.65,123.80) --
	(233.65,123.80) --
	(232.79,124.98) --
	(231.96,126.16) --
	(231.71,126.51) --
	(231.14,127.34) --
	(230.35,128.52) --
	(229.77,129.39) --
	(229.58,129.70) --
	(228.82,130.87) --
	(228.09,132.05) --
	(227.84,132.47) --
	(227.37,133.23) --
	(226.67,134.41) --
	(225.99,135.59) --
	(225.90,135.75) --
	(225.33,136.77) --
	(224.67,137.95) --
	(224.04,139.12) --
	(223.96,139.27) --
	(223.42,140.30) --
	(222.81,141.48) --
	(222.21,142.66) --
	(222.03,143.04) --
	(221.63,143.84) --
	(221.06,145.02) --
	(220.51,146.20) --
	(220.09,147.10) --
	(219.96,147.38) --
	(219.43,148.55) --
	(218.91,149.73) --
	(218.39,150.91) --
	(218.15,151.48) --
	(217.89,152.09) --
	(217.40,153.27) --
	(216.92,154.45) --
	(216.45,155.63) --
	(216.22,156.22) --
	(215.98,156.81) --
	(215.53,157.98) --
	(215.08,159.16) --
	(214.65,160.34) --
	(214.28,161.36) --
	(214.22,161.52) --
	(213.80,162.70) --
	(213.38,163.88) --
	(212.98,165.06) --
	(212.58,166.23) --
	(212.34,166.96) --
	(212.19,167.41) --
	(211.81,168.59) --
	(211.43,169.77) --
	(211.06,170.95) --
	(210.69,172.13) --
	(210.41,173.08) --
	(210.34,173.31) --
	(209.98,174.49) --
	(209.64,175.66) --
	(209.29,176.84) --
	(208.96,178.02) --
	(208.63,179.20) --
	(208.47,179.79) --
	(208.31,180.38) --
	(207.99,181.56) --
	(207.67,182.74) --
	(207.36,183.91) --
	(207.06,185.09) --
	(206.76,186.27) --
	(206.53,187.19) --
	(206.47,187.45) --
	(206.18,188.63) --
	(205.89,189.81) --
	(205.61,190.99) --
	(205.33,192.17) --
	(205.06,193.34) --
	(204.79,194.52) --
	(204.59,195.39) --
	(204.53,195.70) --
	(204.26,196.88) --
	(204.01,198.06) --
	(203.75,199.24) --
	(203.50,200.42) --
	(203.25,201.60) --
	(203.01,202.77) --
	(202.77,203.95) --
	(202.66,204.53) --
	(202.54,205.13) --
	(202.30,206.31) --
	(202.07,207.49) --
	(201.84,208.67) --
	(201.62,209.85) --
	(201.40,211.02) --
	(201.18,212.20) --
	(200.97,213.38) --
	(200.76,214.56) --
	(200.72,214.76) --
	(200.55,215.74) --
	(200.34,216.92) --
	(200.13,218.10) --
	(199.93,219.28) --
	(199.73,220.45) --
	(199.54,221.63) --
	(199.34,222.81) --
	(199.15,223.99) --
	(198.96,225.17) --
	(198.78,226.30) --
	(198.78,226.35) --
	(198.59,227.53) --
	(198.41,228.70) --
	(198.23,229.88) --
	(198.05,231.06) --
	(197.87,232.24) --
	(197.70,233.42) --
	(197.53,234.60) --
	(197.36,235.78) --
	(197.19,236.96) --
	(197.03,238.13) --
	(196.86,239.31) --
	(196.85,239.42) --
	(196.70,240.49) --
	(196.54,241.67) --
	(196.38,242.85) --
	(196.22,244.03) --
	(196.07,245.21) --
	(195.91,246.39) --
	(195.76,247.56) --
	(195.61,248.74) --
	(195.46,249.92) --
	(195.32,251.10) --
	(195.17,252.28) --
	(195.03,253.46) --
	(194.91,254.47) --
	(194.89,254.64) --
	(194.75,255.81) --
	(194.61,256.99) --
	(194.47,258.17) --
	(194.33,259.35) --
	(194.20,260.53) --
	(194.07,261.71) --
	(193.93,262.89) --
	(193.80,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,136.49) --
	(301.02,136.77) --
	(299.50,137.79) --
	(299.27,137.95) --
	(297.58,139.12) --
	(297.56,139.13) --
	(295.93,140.30) --
	(295.62,140.52) --
	(294.32,141.48) --
	(293.69,141.96) --
	(292.76,142.66) --
	(291.75,143.44) --
	(291.24,143.84) --
	(289.81,144.98) --
	(289.76,145.02) --
	(288.32,146.20) --
	(287.88,146.56) --
	(286.91,147.38) --
	(285.94,148.21) --
	(285.54,148.55) --
	(284.21,149.73) --
	(284.00,149.91) --
	(282.90,150.91) --
	(282.07,151.68) --
	(281.62,152.09) --
	(280.38,153.27) --
	(280.13,153.51) --
	(279.16,154.45) --
	(278.19,155.41) --
	(277.97,155.63) --
	(276.81,156.81) --
	(276.26,157.38) --
	(275.68,157.98) --
	(274.57,159.16) --
	(274.32,159.43) --
	(273.48,160.34) --
	(272.41,161.52) --
	(272.38,161.55) --
	(271.37,162.70) --
	(270.45,163.77) --
	(270.35,163.88) --
	(269.35,165.06) --
	(268.51,166.07) --
	(268.37,166.23) --
	(267.41,167.41) --
	(266.57,168.47) --
	(266.47,168.59) --
	(265.55,169.77) --
	(264.65,170.95) --
	(264.64,170.97) --
	(263.76,172.13) --
	(262.89,173.31) --
	(262.70,173.57) --
	(262.04,174.49) --
	(261.20,175.66) --
	(260.76,176.29) --
	(260.38,176.84) --
	(259.57,178.02) --
	(258.83,179.14) --
	(258.78,179.20) --
	(258.00,180.38) --
	(257.24,181.56) --
	(256.89,182.11) --
	(256.49,182.74) --
	(255.75,183.91) --
	(255.03,185.09) --
	(254.95,185.22) --
	(254.31,186.27) --
	(253.61,187.45) --
	(253.02,188.47) --
	(252.92,188.63) --
	(252.24,189.81) --
	(251.58,190.99) --
	(251.08,191.88) --
	(250.92,192.17) --
	(250.28,193.34) --
	(249.64,194.52) --
	(249.14,195.47) --
	(249.02,195.70) --
	(248.40,196.88) --
	(247.80,198.06) --
	(247.20,199.23) --
	(247.20,199.24) --
	(246.61,200.42) --
	(246.03,201.60) --
	(245.47,202.77) --
	(245.27,203.19) --
	(244.90,203.95) --
	(244.35,205.13) --
	(243.81,206.31) --
	(243.33,207.36) --
	(243.27,207.49) --
	(242.74,208.67) --
	(242.22,209.85) --
	(241.71,211.02) --
	(241.39,211.76) --
	(241.20,212.20) --
	(240.70,213.38) --
	(240.21,214.56) --
	(239.72,215.74) --
	(239.46,216.40) --
	(239.25,216.92) --
	(238.77,218.10) --
	(238.31,219.28) --
	(237.85,220.45) --
	(237.52,221.30) --
	(237.39,221.63) --
	(236.94,222.81) --
	(236.50,223.99) --
	(236.07,225.17) --
	(235.64,226.35) --
	(235.58,226.50) --
	(235.21,227.53) --
	(234.79,228.70) --
	(234.38,229.88) --
	(233.97,231.06) --
	(233.65,232.01) --
	(233.57,232.24) --
	(233.17,233.42) --
	(232.77,234.60) --
	(232.39,235.78) --
	(232.00,236.96) --
	(231.71,237.87) --
	(231.62,238.13) --
	(231.25,239.31) --
	(230.88,240.49) --
	(230.51,241.67) --
	(230.15,242.85) --
	(229.79,244.03) --
	(229.77,244.10) --
	(229.44,245.21) --
	(229.09,246.39) --
	(228.74,247.56) --
	(228.40,248.74) --
	(228.07,249.92) --
	(227.84,250.74) --
	(227.74,251.10) --
	(227.40,252.28) --
	(227.08,253.46) --
	(226.75,254.64) --
	(226.44,255.81) --
	(226.12,256.99) --
	(225.90,257.84) --
	(225.81,258.17) --
	(225.50,259.35) --
	(225.20,260.53) --
	(224.89,261.71) --
	(224.60,262.89) --
	(224.30,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,179.14) --
	(301.37,179.20) --
	(300.11,180.38) --
	(299.50,180.96) --
	(298.88,181.56) --
	(297.68,182.74) --
	(297.56,182.85) --
	(296.50,183.91) --
	(295.62,184.80) --
	(295.34,185.09) --
	(294.20,186.27) --
	(293.69,186.81) --
	(293.09,187.45) --
	(291.99,188.63) --
	(291.75,188.89) --
	(290.92,189.81) --
	(289.87,190.99) --
	(289.81,191.05) --
	(288.83,192.17) --
	(287.88,193.28) --
	(287.82,193.34) --
	(286.82,194.52) --
	(285.94,195.58) --
	(285.84,195.70) --
	(284.88,196.88) --
	(284.00,197.97) --
	(283.94,198.06) --
	(283.01,199.24) --
	(282.09,200.42) --
	(282.07,200.45) --
	(281.19,201.60) --
	(280.31,202.77) --
	(280.13,203.02) --
	(279.44,203.95) --
	(278.59,205.13) --
	(278.19,205.68) --
	(277.75,206.31) --
	(276.92,207.49) --
	(276.26,208.45) --
	(276.11,208.67) --
	(275.30,209.85) --
	(274.52,211.02) --
	(274.32,211.32) --
	(273.74,212.20) --
	(272.97,213.38) --
	(272.38,214.31) --
	(272.22,214.56) --
	(271.48,215.74) --
	(270.75,216.92) --
	(270.45,217.41) --
	(270.03,218.10) --
	(269.32,219.28) --
	(268.62,220.45) --
	(268.51,220.64) --
	(267.93,221.63) --
	(267.25,222.81) --
	(266.58,223.99) --
	(266.57,224.01) --
	(265.92,225.17) --
	(265.27,226.35) --
	(264.64,227.52) --
	(264.63,227.53) --
	(263.99,228.70) --
	(263.37,229.88) --
	(262.76,231.06) --
	(262.70,231.17) --
	(262.15,232.24) --
	(261.55,233.42) --
	(260.96,234.60) --
	(260.76,234.99) --
	(260.37,235.78) --
	(259.79,236.96) --
	(259.23,238.13) --
	(258.83,238.98) --
	(258.67,239.31) --
	(258.11,240.49) --
	(257.56,241.67) --
	(257.02,242.85) --
	(256.89,243.14) --
	(256.49,244.03) --
	(255.96,245.21) --
	(255.44,246.39) --
	(254.95,247.51) --
	(254.93,247.56) --
	(254.42,248.74) --
	(253.91,249.92) --
	(253.42,251.10) --
	(253.02,252.08) --
	(252.93,252.28) --
	(252.44,253.46) --
	(251.96,254.64) --
	(251.49,255.81) --
	(251.08,256.87) --
	(251.03,256.99) --
	(250.56,258.17) --
	(250.10,259.35) --
	(249.65,260.53) --
	(249.21,261.71) --
	(249.14,261.89) --
	(248.77,262.89) --
	(248.33,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,212.31) --
	(300.50,213.38) --
	(299.50,214.55) --
	(299.49,214.56) --
	(298.49,215.74) --
	(297.56,216.85) --
	(297.51,216.92) --
	(296.54,218.10) --
	(295.62,219.24) --
	(295.59,219.28) --
	(294.66,220.45) --
	(293.74,221.63) --
	(293.69,221.70) --
	(292.83,222.81) --
	(291.94,223.99) --
	(291.75,224.25) --
	(291.06,225.17) --
	(290.20,226.35) --
	(289.81,226.88) --
	(289.35,227.53) --
	(288.51,228.70) --
	(287.88,229.61) --
	(287.68,229.88) --
	(286.87,231.06) --
	(286.07,232.24) --
	(285.94,232.43) --
	(285.28,233.42) --
	(284.50,234.60) --
	(284.00,235.35) --
	(283.73,235.78) --
	(282.97,236.96) --
	(282.22,238.13) --
	(282.07,238.38) --
	(281.49,239.31) --
	(280.76,240.49) --
	(280.13,241.53) --
	(280.04,241.67) --
	(279.33,242.85) --
	(278.64,244.03) --
	(278.19,244.79) --
	(277.95,245.21) --
	(277.27,246.39) --
	(276.60,247.56) --
	(276.26,248.17) --
	(275.93,248.74) --
	(275.28,249.92) --
	(274.64,251.10) --
	(274.32,251.69) --
	(274.00,252.28) --
	(273.37,253.46) --
	(272.75,254.64) --
	(272.38,255.34) --
	(272.13,255.81) --
	(271.53,256.99) --
	(270.93,258.17) --
	(270.45,259.14) --
	(270.34,259.35) --
	(269.75,260.53) --
	(269.17,261.71) --
	(268.61,262.89) --
	(268.51,263.09) --
	(268.04,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,238.85) --
	(301.08,239.31) --
	(300.18,240.49) --
	(299.50,241.41) --
	(299.30,241.67) --
	(298.43,242.85) --
	(297.58,244.03) --
	(297.56,244.05) --
	(296.73,245.21) --
	(295.90,246.39) --
	(295.62,246.78) --
	(295.08,247.56) --
	(294.27,248.74) --
	(293.69,249.61) --
	(293.48,249.92) --
	(292.69,251.10) --
	(291.91,252.28) --
	(291.75,252.53) --
	(291.15,253.46) --
	(290.39,254.64) --
	(289.81,255.55) --
	(289.65,255.81) --
	(288.91,256.99) --
	(288.18,258.17) --
	(287.88,258.67) --
	(287.46,259.35) --
	(286.75,260.53) --
	(286.06,261.71) --
	(285.94,261.91) --
	(285.36,262.89) --
	(284.68,264.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] (301.44,260.56) --
	(300.64,261.71) --
	(299.84,262.89) --
	(299.50,263.39) --
	(299.04,264.07);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.16, 30.69) rectangle (302.65,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 27.66) {0.0};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 51.23) {0.1};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 74.80) {0.2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21, 98.38) {0.3};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,121.95) {0.4};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,145.52) {0.5};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,169.10) {0.6};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,192.67) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,216.25) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,239.82) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 29.21,263.39) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 30.69) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 54.26) --
	( 34.16, 54.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41, 77.83) --
	( 34.16, 77.83);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,101.41) --
	( 34.16,101.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,124.98) --
	( 34.16,124.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,148.55) --
	( 34.16,148.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,172.13) --
	( 34.16,172.13);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,195.70) --
	( 34.16,195.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,219.28) --
	( 34.16,219.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,242.85) --
	( 34.16,242.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.41,266.42) --
	( 34.16,266.42);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.16, 27.94) --
	( 34.16, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.41, 27.94) --
	(168.41, 30.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (302.65, 27.94) --
	(302.65, 30.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 34.16, 19.68) {0.5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (168.41, 19.68) {1.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (302.65, 19.68) {2.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (168.41,  7.64) {Relative mean asymptomatic generation interval, $G_a/G_s$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,148.55) {Intrinsic proportion of asymptomatic transmission, $z$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (313.65, 28.75) rectangle (355.85,268.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\node[inner sep=0pt,outer sep=0pt,anchor=south west,rotate=  0.00] at (319.15,  34.25) {
	\pgfimage[width= 14.45pt,height=213.40pt,interpolate=true]{figheatmap_R0_03_ras2}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11, 36.55) {1.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11, 81.60) {2.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,126.65) {3.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,171.70) {4.5};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.88] at (339.11,216.75) {5.5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at (319.15,254.21) {$\mathcal{R}_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15, 39.58) -- (322.05, 39.58);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15, 84.63) -- (322.05, 84.63);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,129.68) -- (322.05,129.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,174.73) -- (322.05,174.73);

\path[draw=drawColor,line width= 0.2pt,line join=round] (319.15,219.78) -- (322.05,219.78);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72, 39.58) -- (333.61, 39.58);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72, 84.63) -- (333.61, 84.63);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,129.68) -- (333.61,129.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,174.73) -- (333.61,174.73);

\path[draw=drawColor,line width= 0.2pt,line join=round] (330.72,219.78) -- (333.61,219.78);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.32] at ( 34.16,274.49) {B. Basic reproduction number};
\end{scope}
\end{tikzpicture}

\end{document}