### 2020-03-11 acknowledge.tex

\mbox{}\\
\noindent
\textbf{Acknowledgments:} The authors thank John Glasser for comments
and discussion on the manuscript, particularly on short notice.

\noindent
\textbf{Data Availabililty:} All code is available at https://github.com/mac-theobio/coronavirus\_asymptomatic.