### 2020-04-05 acknowledge.tex
done?

\mbox{}\\
\noindent
\textbf{Acknowledgments:} The authors thank John Glasser for comments
and discussion on the manuscript, particularly on short notice. 
The authors thank multiple reviewers for their feedback.
Research effort by JSW was enabled
by support from grants from the Simons Foundation (SCOPE Award ID 329108), the Army Research
Office (W911NF1910384), National Institutes of Health (1R01AI46592-01),
and National Science Foundation (1806606 and 1829636). JD was supported, in part, by
grants from the Canadian Institutes of Health Research.

\noindent
\textbf{Data Availabililty:} All code is available at \url{https://github.com/mac-theobio/coronavirus_asymptomatic}.